---
title: "ZQM Data Exploration"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: material
    highlight: github
    df_print: kable
---

```{r setup, include=FALSE}
library(readxl)
library(terra)
library(sf)
library(tidyverse)
library(ggrepel)
library(ggpubr)
library(rmdformats)
library(leaflet)
library(ggspatial)
library(ggthemes)

my_opts = read_csv(paste0(here::here(),"/Options.csv"))

knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_knit$set(root.dir = paste0(my_opts$zqm_operations_data_folder,"Watercraft Inspection Data/Multiyear data/"))

outlier.multiple = 25
```

This R Markdown document explores the datasets we plan to use to inform our assessment of ZQM invasion risk for waterbodies in BC.
  
```{r import_waterbody_data_and_filtering}
# Inspections joined to water bodies.

waterb = read_sf(paste0(my_opts$base_dir,"03_PrioritizationModel/data/waterb_with_data.gpkg"))

#Remove average.angler.days
waterb$average.angler.days = NULL
waterb$WB.access = NULL
```

```{r ensure_water_temp_vars_are_factors}
waterb = waterb |> 
  dplyr::mutate(dplyr::across(dplyr::starts_with("water_"), \(x) factor(x, levels = c("No Risk","Low Risk","Moderate Risk","High Risk"))))
```

## Filter Tracking
When cleaning the inspection data, I removed records that were headed to the ocean or dry storage. Some inspection records also did not match to any waterbody. 

```{r inspection_filter_trackers}
read_excel(paste0(my_opts$zqm_operations_data_folder,"Watercraft Inspection Data/Multiyear data/Inspection_filter_tracker_with_delta.xlsx")) %>% as_tibble(.)
```

And it is important to note the number of waterbodies included in this analysis, as well as the number of waterbodies for which we have watercraft inspection data.

```{r filter_steps}
read_excel(paste0(my_opts$zqm_operations_data_folder,"Watercraft Inspection Data/Multiyear data/waterbody_filtering_steps.xlsx")) %>% 
  as_tibble(.)
```

```{r data_exploration_functions}

bc = read_sf(paste0(my_opts$remote_spatial_data,"shared_data_sets/bc_shapefile.shp"))

options(scipen = 999)

exploration_big10 = function(x){
  dat = waterb[!is.na(waterb[[x]]),] 
  
  dat %>% 
    ungroup() %>% 
    st_drop_geometry() %>% 
    dplyr::arrange(desc(x)) %>% 
    slice(1:10) %>% 
    mutate(WaterbodyName = coalesce(GNIS_NA, as.character(WATERBO))) %>% 
    summarise(Top10 = str_c(WaterbodyName,collapse = ", "))
}

exploration_battery = function(x){
  
  dat = waterb %>% ungroup() %>% 
    filter(!is.na(.[[x]])) %>% 
    filter(.[[x]] > 0)
  
  #Find outliers.
  outliers = dat %>% 
    mutate(mean_value = mean(.[[x]],na.rm=T)) %>% 
    filter(.[[x]] > mean_value*outlier.multiple)
  
  #Grab the name of the variable of interest.
  data.name = x
  
  #Make a simple histogram.
  p1 = ggplot(dat) + 
    geom_histogram(aes(.data[[x]]))

  #Make a boxplot with violin plot overlay that highlights the outlier points with labels.
  p2 = ggplot() + 
    geom_boxplot(data=dat,aes(y=.data[[x]], x = data.name)) + 
    geom_violin(data=dat,aes(y=.data[[x]], x = data.name), alpha=0.6) +
    labs(x = "",y=data.name) +
    {if(length(outliers$WATERSH) > 0)geom_point(data = outliers, aes(y=.data[[x]], x = data.name), col = "red")}+ 
    {if(length(outliers$WATERSH) > 0)geom_label_repel(data = outliers, aes(y=.data[[x]], x = data.name,label=coalesce(.data$GNIS_NA,as.character(.data$WATERBO))),
                     force = 100, max.overlaps = 20)}
 
  # Give an idea of the spatial coverage we have in the province.
  if(x != "PopPressure") {
    # browser()
  p3 = ggplot() + 
    geom_sf(data = bc, fill = "antiquewhite", col = "transparent") +
    geom_sf(data = dat, aes(col = .data[[x]], fill = .data[[x]])) + 
    scale_fill_gradient(low = "green", high = "red") +
    scale_color_gradient(low = "green", high = "red") +
    guides(col = "none") +
    # labs(fill = stringr::str_replace_all(.data[[x]],"_","\n"),
    #      col = stringr::str_replace_all(.data[[x]],"_","\n")) +
    ggtitle("Data coverage across the province",
            subtitle = paste0("variable: ",x)) + 
    ggthemes::theme_map()
  }
  if(x == "PopPressure") {
    p3 = ggplot() + 
      geom_sf(data = bc, fill = "darkgreen")
  }
  
  return(ggarrange(p3,ggarrange(p1,p2,nrow=2),ncol=2))
}


exploration_battery_categorical = function(x){
  
  dat = waterb[!is.na(waterb[[x]]),]
  
  p1 = ggplot() + 
  geom_bar(data=dat, aes(.data[[x]]))

  # Give an idea of the spatial coverage we have in the province.
  p2 = ggplot() + 
    geom_sf(data = bc, fill = "antiquewhite", col = "transparent") +
    geom_sf(data = dat, aes(col = .data[[x]])) + 
    ggtitle("Data coverage across the province",
            subtitle = paste0("variable: ",x))
  
  return(list(p1,p2))
}

my.summary = function(x){
  my.sum = as.data.frame(as.matrix(unlist(summary(x))))

  my.sum = my.sum %>% 
    mutate(variable = row.names(my.sum)) %>% 
    filter(variable %in% c("Min.","Median","Mean","Max.")) %>% 
    pivot_wider(names_from = variable, values_from = V1) %>% 
    mutate(Data_Coverage = paste0(round(100*(length(x[!is.na(x)])/length(x)), 3), "%"))
  
  return(my.sum)
}


```

# Data Exploration 

### Total Inspections
```{r}
my.summary(waterb$TotalInspections)

exploration_big10(x = "TotalInspections")

exploration_battery(x = "TotalInspections")
```

### Sum of Days Fished

```{r angling_days}
my.summary(waterb$days_fished_2023_survey)

exploration_big10(x = "days_fished_2023_survey")

exploration_battery(x = "days_fished_2023_survey")
```

### Campgrounds

These data are a combination of Sam Fischer _et al._'s data and angler survey data.

```{r campgrounds}
my.summary(waterb$Campgrounds)

exploration_big10(x = "Campgrounds")

exploration_battery(x = "Campgrounds")
```

### Facilities

"Facilities" refers to the number of touristic facilities (public toilets, tourist information, viewpoints, parks, attractions, and picnic sites) within 500 meters of the water body. These data come from the PhD work of Sam Fischer _et al._.

```{r facilites}
my.summary(waterb$Facilities)

exploration_big10(x = "Facilities")

exploration_battery(x = "Facilities")
```

### Marinas

These data are a combination of Sam Fischer _et al._'s data and angler survey data.

```{r marinas}
my.summary(waterb$Marinas)

exploration_big10(x = "Marinas")

exploration_battery(x = "Marinas")
```

### Dams

We have data from BC Hydro and Fortis on the location and capacity (in megawatts) of dams throughout British Columbia. To find the most likely spatial match for each dam, I used only large water bodies (>= 3,000,000 square meters). Without this filter, the nearest water body tended to be very small ponds right beside the location of the dam.

### SummedDamCapacity
```{r summed_dam_capacity}
my.summary(waterb$SummedDamCapacity)

waterb = waterb %>% 
  mutate(NumberDams = as.numeric(NumberDams),
         SummedDamCapacity = as.numeric(SummedDamCapacity))

exploration_big10(x = "SummedDamCapacity")

exploration_battery(x = "SummedDamCapacity")

dams_sf = read_sf(paste0(my_opts$remote_spatial_data,"shared_data_sets/dams_summarised_sf.shp")) %>% st_transform(crs = 4326)

waterb_with_dams = waterb %>% 
  filter(NumberDams > 0 )%>% 
  st_transform(crs = 4326)

dams_sf = dams_sf %>% 
  mutate(DC_bin = as.numeric(cut(DmCpcty, breaks = 
                          
                           BAMMtools::getJenksBreaks(dams_sf$DmCpcty, k = 4)))) %>% 
  mutate(DC_bin = replace_na(DC_bin, 1))

icon.blue <- makeAwesomeIcon(markerColor = 'blue', iconColor = 'blue')
icon.red <- makeAwesomeIcon(markerColor = 'red', iconColor = 'red')
icon.green <- makeAwesomeIcon(markerColor = 'green', iconColor = 'green')

leaflet() %>%
  addTiles() %>%
  addAwesomeMarkers(data = dams_sf[dams_sf$DC_bin==1,], popup = ~Name, label = ~DmCpcty, icon = icon.green) %>%
  addAwesomeMarkers(data = dams_sf[dams_sf$DC_bin==2,], popup = ~Name, label = ~DmCpcty, icon = icon.blue) %>%
  addAwesomeMarkers(data = dams_sf[dams_sf$DC_bin==3,], popup = ~Name, label = ~DmCpcty, icon = icon.red) %>%
  addPolygons(data = waterb_with_dams, color = "red", fillColor = "red", label = ~GNIS_NA)

```

### Number Dams
```{r summed_dam_number}
my.summary(waterb$NumberDams)

exploration_big10(x = "NumberDams")

exploration_battery(x = "NumberDams")

``` 

### Boat Launches

We add together boat launch data from the FFSBC (Freshwatery Fisheries Society of British Columbia), 2010 angler survey data, and the recreation layer in the BCG Warehouse. Even still, we are missing boat launch data for many of the largest, most visited lakes...

```{r boat_launches}
waterb$BoatLaunches = waterb$Boat_Launches_N + waterb$BoatLaunchRecCounter

my.summary(waterb$BoatLaunches)

#Top 20 lakes in terms of size - how do their boat launches look?
waterb %>% 
  mutate(Area = as.numeric(st_area(.))) %>% 
  arrange(desc(Area)) %>% 
  slice(1:20) %>% 
  st_drop_geometry() %>% 
  select(GNIS_NA, WATERBO, Boat_Launches_N)

exploration_big10(x = "BoatLaunches")

exploration_battery(x = "BoatLaunches")
```

### Dissolved Calcium
```{r}
calc = read_sf(paste0(my_opts$base_dir,'01_DataCleaning/output/ZQM_RiskAssessment_ssp370_10C_threshold_pH_Adjustment_Subwatershed.gpkg'))

calc = calc |> 
  dplyr::mutate(calc_bin = factor(calc_bin,
                                  levels = c(
                                    "No Risk (< 8 mg/L)",
                                    "Low (8 to <15 mg/L)",
                                    "Moderate (15 to <30 mg/L)",
                                    "High (30+ mg/L)",
                                    "No Data"
                                  )))

ggplot() + 
  geom_sf(data = calc, fill = 'grey') + 
  geom_sf(data = calc, aes(fill = calc_bin)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk (< 8 mg/L)" = "green",
                               "Low (8 to <15 mg/L)" = "yellow",
                               "Moderate (15 to <30 mg/L)" = "orange",
                               "High (30+ mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Binned Calcium at the Subwatershed scale',
       subtitle = paste0(nrow(calc[calc$calc_bin != 'No Data',]), ' of 246 subwatersheds with data'))


  
```

### Water Temperature

```{r}
waterb[!is.na(waterb$TotalInspections),] |> sf::st_drop_geometry() |> dplyr::count(water_temp_interp)

waterb_for_plot = waterb |> 
  dplyr::filter(!is.na(TotalInspections),
                !GNIS_NA %in% c("Dry Storage","Pacific Ocean"))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = waterb_for_plot, fill = 'grey') + 
  geom_sf(data = waterb_for_plot, aes(fill = water_temp_interp, col = water_temp_interp)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  scale_colour_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Water Temperature',
       color = 'Binned Water Temperature',
       title = 'Binned Water Temperature at the Subwatershed scale',
       subtitle = paste0(nrow(waterb_for_plot[waterb_for_plot$water_temp_interp != 'No Data',]), ' of 246 subwatersheds with data'))
```

### Water Chlorophyll

```{r}
waterb[!is.na(waterb$TotalInspections),] |> sf::st_drop_geometry() |> dplyr::count(water_chlor_interp)

waterb_for_plot = waterb |> 
  dplyr::filter(!is.na(TotalInspections),
                !GNIS_NA %in% c("Dry Storage","Pacific Ocean"))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = waterb_for_plot, fill = 'grey') + 
  geom_sf(data = waterb_for_plot, aes(fill = water_chlor_interp, col = water_chlor_interp)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  scale_colour_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Water \nChlorophyll',
       color = 'Binned Water \nChlorophyll',
       title = 'Binned Water Chlorophyll at the Subwatershed scale',
       subtitle = paste0(nrow(waterb_for_plot[waterb_for_plot$water_chlor_interp != 'No Data',]), ' of 246 subwatersheds with data'))
```

### Water Dissolved Oxygen

```{r}
waterb[!is.na(waterb$TotalInspections),] |> sf::st_drop_geometry() |> dplyr::count(water_dis_ox_interp)

waterb_for_plot = waterb |> 
  dplyr::filter(!is.na(TotalInspections),
                !GNIS_NA %in% c("Dry Storage","Pacific Ocean"))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = waterb_for_plot, fill = 'grey') + 
  geom_sf(data = waterb_for_plot, aes(fill = water_dis_ox_interp, col = water_dis_ox_interp)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  scale_colour_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Water \nDissolved Oxygen',
       color = 'Binned Water \nDissolved Oxygen',
       title = 'Binned Water Dissolved Oxygen at the Subwatershed scale',
       subtitle = paste0(nrow(waterb_for_plot[waterb_for_plot$water_dis_ox_interp != 'No Data',]), ' of 246 subwatersheds with data'))
```

### Water Total Phosphorous

```{r}
waterb[!is.na(waterb$TotalInspections),] |> sf::st_drop_geometry() |> dplyr::count(water_tot_phos_interp)

waterb_for_plot = waterb |> 
  dplyr::filter(!is.na(TotalInspections),
                !GNIS_NA %in% c("Dry Storage","Pacific Ocean"))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = waterb_for_plot, fill = 'grey') + 
  geom_sf(data = waterb_for_plot, aes(fill = water_tot_phos_interp, col = water_tot_phos_interp)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  scale_colour_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Water \nPhosphorous',
       color = 'Binned Water \nPhosphorous',
       title = 'Binned Water Phosphorous at the Subwatershed scale',
       subtitle = paste0(nrow(waterb_for_plot[waterb_for_plot$water_tot_phos_interp != 'No Data',]), ' of 246 subwatersheds with data'))
```

### Water pH

```{r}
waterb[!is.na(waterb$TotalInspections),] |> sf::st_drop_geometry() |> dplyr::count(water_pH)

waterb_for_plot = waterb |> 
  dplyr::filter(!is.na(TotalInspections),
                !GNIS_NA %in% c("Dry Storage","Pacific Ocean"))

ggplot() + 
  geom_sf(data = bc) + 
  geom_sf(data = waterb_for_plot, fill = 'grey') + 
  geom_sf(data = waterb_for_plot, aes(fill = water_pH, col = water_pH)) + 
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  scale_colour_manual(values = c("No Data" = "grey",
                               "No Risk" = "green",
                               "Low Risk" = "yellow",
                               "Moderate Risk" = "orange",
                               "High Risk" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Water pH',
       color = 'Binned Water pH',
       title = 'Binned Water pH at the Subwatershed scale',
       subtitle = paste0(nrow(waterb_for_plot[waterb_for_plot$water_pH != 'No Data',]), ' of 246 subwatersheds with data'))
```

### Population Density

Population density is calculated at the subwatershed scale.

```{r}
my.summary(waterb$population_density)

exploration_big10(x = "population_density")

exploration_battery(x = "population_density")
```

### Watercraft Operation Restrictions

These data come from Transport Canada and indicate lakes in which one may not operate any vessel, may not operate power vessels, or may not operate power/electric vessels.

```{r waterbody_restrictions}
waterb %>% 
  st_drop_geometry() %>% 
  group_by(OperRes) %>% 
  summarise(Number = n()) %>% 
  pivot_wider(names_from = OperRes, values_from = Number) %>% 
  mutate(Data_Coverage = paste0(round(100*(nrow(waterb)-`NA`)/nrow(waterb), 3), "%")) %>% 
  select(-`NA`)

results = exploration_battery_categorical(x = "OperRes")

results[[1]]

results[[2]]
```

### DFO Predicted Suitability for Zebra and Quagga Mussels (Ca2+ model) {.tabset}

These interpolated maps utilized EMS data for BC (2000-2020), filtered away any negative or 0 values for calcium, applied a spatial filter to remove calcium points outside of waterbodies, and removed calcium values of type 'Industrial Effluent.' The predicted value is the establishment potential based on (dissolved) calcium. There are also MaxEnt-derived models that give predictions of Zebra and Quagga mussel establishment potential (based on the following predictor variables: Ca2+, Max Temp in warmest month, min Temp in coldest month, min Temp in wettest quarter, precipitation in wettest month, precip seasonality, precip in coldest quarter, and pH, the last of which was by far the least significant predictor!) but those models' predictions perhaps under-predict the suitability of habitat across North America, so they have not been utilized or visualized here.

```{r}
# DFO interpolated calcium TIF
dfo_calc_qm = terra::rast(paste0(my_opts$base_dir,'03_PrioritizationModel/data/DFO_2024_QM_Calcium_Establishment_Potential.tif'))
dfo_calc_zm = terra::rast(paste0(my_opts$base_dir,'03_PrioritizationModel/data/DFO_2024_ZM_Calcium_Establishment_Potential.tif'))

# Clip to BC.
bc_vect = terra::project(terra::vect(dplyr::summarise(bc)), terra::crs(dfo_calc_qm))

dfo_calc_qm = terra::mask(terra::crop(dfo_calc_qm, bc_vect), bc_vect)
dfo_calc_zm = terra::mask(terra::crop(dfo_calc_zm, bc_vect), bc_vect)

# Take larger predicted value for any given raster cell.
dfo_calc = max(dfo_calc_qm, dfo_calc_zm,na.rm=T)

terra::writeRaster(dfo_calc, paste0(my_opts$base_dir,'03_PrioritizationModel/data/DFO_2024_ZQM_Calcium_Establishment_Potential.tif'), overwrite = T)

# DFO subwatershed values.
dfo_qm_subdr = sf::read_sf(paste0(my_opts$base_dir,'03_PrioritizationModel/data/DFO_2024_ZM_QM_Shapefiles/QM_Ecological_Risk_Sub-Drainage.shp'))
dfo_zm_subdr = sf::read_sf(paste0(my_opts$base_dir,'03_PrioritizationModel/data/DFO_2024_ZM_QM_Shapefiles/ZM_Ecological_Risk_Sub-Drainage.shp'))

dfo_qm_subdr = sf::st_intersection(dfo_qm_subdr, sf::st_transform(dplyr::summarise(bc), terra::crs(dfo_qm_subdr)))
dfo_zm_subdr = sf::st_intersection(dfo_zm_subdr, sf::st_transform(dplyr::summarise(bc), terra::crs(dfo_zm_subdr)))
```

#### Zebra Mussel Establishment Potential (Ca2+)
```{r}
ggplot() + 
  tidyterra::geom_spatraster(data = terra::project(dfo_calc_zm, 'EPSG:3005')) + 
  scale_fill_distiller(palette = 'Spectral', na.value = 'transparent') + 
  ggthemes::theme_map()
```

#### Quagga Mussel Establishment Potential (Ca2+)

```{r}
ggplot() + 
  tidyterra::geom_spatraster(data = terra::project(dfo_calc_qm, 'EPSG:3005')) + 
  scale_fill_distiller(palette = 'Spectral', na.value = 'transparent') + 
  ggthemes::theme_map()
```

#### Combined 
```{r}
ggplot() + 
  tidyterra::geom_spatraster(data = terra::project(dfo_calc, 'EPSG:3005')) + 
  scale_fill_distiller(palette = 'Spectral', na.value = 'transparent') + 
  ggthemes::theme_map()
```

### DFO Predicted Ecological Risk at Sub-Drainage level {.tabset}

#### Zebra Mussel
```{r}
ggplot() + geom_sf(data = dfo_zm_subdr |> sf::st_transform(3005), aes(fill = ZMME_Mx_E)) + 
  scale_fill_manual(values = c("Low" = "lightgreen", "Moderate" = 'gold', "High" = 'darkred')) + 
  ggthemes::theme_map()
```

#### Quagga Mussel
```{r}
ggplot() + geom_sf(data = dfo_qm_subdr |> sf::st_transform(3005), aes(fill = QMME_Mx_E)) + 
  scale_fill_manual(values = c("Low" = "lightgreen", "Moderate" = 'gold', "High" = 'darkred')) + 
  ggthemes::theme_map()
```

```{r variable_choice_table}
var_table = waterb %>% 
  st_drop_geometry() %>% 
  select(-HR_Non_Motorized_Counter,-HR_Complex_Counter,-HR_Very_Complex_Counter,-HR_Simple_Counter,-LowRisk_Complex_Counter,-LowRisk_Non_Motorized_Counter,-LowRisk_Simple_Counter,-LowRisk_Complex_Counter) %>% 
  mutate(across(everything(), as.character)) %>% 
  pivot_longer(cols = -c(WATERBO,WATERSH,GNIS_NA)) %>% 
  filter(!is.na(value)) %>% 
  #Find the total number of records for each variable.
  group_by(name) %>% 
  mutate(Waterbodies_with_data = paste0(n()," (",round(100*n()/nrow(waterb),2),"%)")) %>% 
  #Find the number of records for each variable / subwatershed combination.
  group_by(WATERSH,name) %>% 
  mutate(Records_at_subwatershed_level = n()) %>% 
  ungroup() %>% 
  select(name, Waterbodies_with_data, Records_at_subwatershed_level) %>% 
  distinct() %>% 
  #Find the number of unique subwatersheds with values
  group_by(name,Waterbodies_with_data) %>% 
  summarise(Subwatersheds_with_data = n()) %>% 
  rename(Variable = name)

# var_table = var_table %>% 
#   ungroup() %>% 
#   add_row(Variable = "PopPressure at subwatershed",
#           Waterbodies_with_data = "424898 (100%)",
#           Subwatersheds_with_data = 246)

var_table = var_table %>% 
  arrange(desc(as.numeric(str_extract(Waterbodies_with_data,"[0-9]*(?= \\()"))))

# var_table$Focus = c("Value","Geography","Access/Use",
#                    "Access/Use","Access/Use",
#                    "Access/Use","Access/Use","Access/Use",
#                    "Access/Use","Access/Use",
#                    "Access/Use","Access/Use",
#                    "Access/Use","Access/Use",
#                    "Value","Value")

var_table

openxlsx::write.xlsx(waterb %>% 
           st_drop_geometry(), "waterb_with_data_explored.xlsx",
           overwrite = T)
```

