---
title: "ZQM Waterbody Sampling Shortlist"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output:
  rmdformats::robobook:
  self_contained: true
highlight: tango
df_print: kable
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(readxl)
library(terra)
library(tidyverse)
library(dplyr)
library(stringr)
library(tidyr)
library(sf)
library(leaflet)
library(BAMMtools)
library(RColorBrewer)
library(openxlsx)
library(diffdf)
library(bcdata)
library(readr)

my_opts = read_csv(paste0(here::here(),"/Options.csv"))

my.year = my_opts$year + 1

knitr::opts_chunk$set(
	echo = FALSE,
	fig.height = 10,
	fig.width = 10,
	message = FALSE,
	warning = FALSE,
	root.dir = paste0(my_opts$zqm_operations_data_folder,"Watercraft Inspection Data/Multiyear data/")
)
# 
# if(interactive()){
#   if(stringr::str_extract(getwd(),"[0-9]{2}_.*") != "03_PrioritizationModel"){
#     setwd(paste0(stringr::str_remove(getwd(),"[0-9]{2}_.*")))
#   }
# }

update_remote_output_files = T

my_path = ifelse(interactive(), "03_PrioritizationModel/","")

source(paste0(my_path,'utility_functions.R'))
```

This R Markdown script takes the output of the script '03_ZQM_stats_and_waterbody output.Rmd' (namely, waterbodies with raw and binned risk estimates for choice variables) to produce a short-list of priority waterbodies vis-a-vis Zebra and Quagga Mussel risk.

```{r read_in_data}
# Read RDS.
dat = read_rds(paste0(my_path,"data/data_for_making_shortlist.rds"))
# if(interactive()){
#   dat = read_rds('03_PrioritizationModel/data/data_for_making_shortlist.rds')
# } else {
# dat = read_rds('data/data_for_making_shortlist.rds')
# }

waterb = dat[[1]]
waterb_risk = dat[[2]]
waterb_og_values = dat[[3]]
subw_w_dat = dat[[4]]
naturalbreaks = dat[[5]]
```

# Filter Waterbodies to get High-priority Short-list

```{r put_aside_mussel_fouled}
waterb_with_any_mf = waterb |> 
  filter(!is.na(NumberMusselFouled) & NumberMusselFouled > 0)

# + 3. Waterbodies with a binned risk estimate of 2 and a subwatershed-scale dissolved calcium level that was the lowest risk were dropped.
```

```{r join_risk_bins_with_og_values}
#Join the OG values for our risk variables to this layer with binned values.
waterb_og_values_and_bins = waterb_risk |> 
  left_join(waterb_og_values |> 
              filter(!is.na(TotalInspections)) |> 
              st_drop_geometry() |> 
              dplyr::rename(
                SummedDamCapacity_unbinned = SummedDamCapacity,
                # distinct_SAR_unbinned = distinct_SAR,
                distinct_Fed_SARA_listed_SAR_unbinned = distinct_Fed_SARA_listed_SAR,
                distinct_CDC_listed_SAR_unbinned = distinct_CDC_listed_SAR,
                OperRes_unbinned = OperRes) |> 
              dplyr::mutate(SummedDamCapacity_unbinned = replace_na(SummedDamCapacity_unbinned, 0),
                            # distinct_SAR_unbinned = replace_na(distinct_SAR_unbinned, 0),
                            distinct_Fed_SARA_listed_SAR_unbinned = replace_na(distinct_Fed_SARA_listed_SAR_unbinned),
                            distinct_CDC_listed_SAR_unbinned = replace_na(distinct_CDC_listed_SAR_unbinned),
                            OperRes_unbinned = replace_na(OperRes_unbinned, 'No Restriction'))
  ) #|> 
 # left_join(subw_w_dat |> 
             # st_drop_geometry() |> 
             # dplyr::select(WATERSH = WATERSHED_,
             #               distinct_SAR_value = distinct_SAR,
             #               SummedDamCapacity_value = SummedDamCapacity))

#Write out to W: drive.
if(update_remote_output_files){
  if(!dir.exists(paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/"))){
    dir.create(paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/"),
               recursive = T)
  }
  sf::write_sf(waterb_og_values_and_bins,
           paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/Waterbodies_with_binned_and_original_values.gpkg"))
}
```

```{r drop_duplicated_rivers}
#Find the river polygons that actually belong to the same water bodies. n = 39.
duplicated_river_names = waterb_risk |> 
                     filter(str_detect(GNIS_NA, "River"),
                            GNIS_NA != "Elk River") |> 
                     filter(duplicated(GNIS_NA)) |> 
                     st_drop_geometry() |> 
                     dplyr::select(GNIS_NA)

duplicated_rivers = waterb_risk |> 
 filter(GNIS_NA %in% all_of(duplicated_river_names$GNIS_NA)) |> 
 dplyr::select(WATERSH,WATERBO,GNIS_NA)

waterb_risk_rivers_to_merge = waterb_risk |> 
 filter(paste0(WATERSH,WATERBO,GNIS_NA) %in% 
          all_of(paste0(duplicated_rivers$WATERSH,
                        duplicated_rivers$WATERBO,
                        duplicated_rivers$GNIS_NA)))
 
# Also double check that the rivers we're looking at are touching each other. Or, look at how far apart river pieces with the same name are.
# If they are closer than the length of either river, join them up.
unique_duplicated_river_names = unique(waterb_risk_rivers_to_merge$GNIS_NA)

waterb_risk_rivers_merged = list()

for(i in 1:length(unique_duplicated_river_names)){

  if(interactive()) print(paste0(i, ' out of ',length(unique_duplicated_river_names)))
  
  the_name = unique_duplicated_river_names[i]
  
  the_rivers = waterb_risk_rivers_to_merge |> 
    dplyr::filter(GNIS_NA == the_name)
  
  if(the_name %in% c('Fraser River','Columbia River','Kootenay River')){
    
    new_rivers = the_rivers |> 
      group_by(GNIS_NA) |> 
      arrange(GNIS_NA,desc(as.numeric(InvasionRisk))) |> 
      summarise(across(!starts_with("geo"), first))
    
    waterb_risk_rivers_merged[[i]] = new_rivers
  } else {
    
    rivers_product = list()
    
    for(y in 1:nrow(the_rivers)){
      river_focus = the_rivers[y,]
      river_length = river_focus |> 
      st_cast("MULTILINESTRING") |> 
      mutate(the_length = st_length(geom)) |> 
      dplyr::pull(the_length)
      
      # Which of the other rivers should we merge?
      rivers_to_merge = st_distance(the_rivers)[y,] <= river_length
      
      new_geom = the_rivers[rivers_to_merge,] |> 
        arrange(GNIS_NA,desc(as.numeric(InvasionRisk))) |> 
         summarise(across(!starts_with("geo"), first))

      rivers_product[[y]] <- new_geom
    }
    
    rivers_products_b = dplyr::bind_rows(
      rivers_product
    )
    # Check to see if some rows are just subsets of other rows, in terms of geometry; remove such subsets...
    subsets_to_remove = st_intersects(rivers_products_b) |> 
      as.data.frame() |> 
      dplyr::filter(row.id != col.id) |> 
      mutate(combo = paste0(order(row.id, col.id), collapse = ',')) |> 
      mutate(is_duplicated = duplicated(combo)) |> 
      pull(is_duplicated)
    
    if(length(subsets_to_remove) > 0){
      rivers_products_b = rivers_products_b[-subsets_to_remove,]
    }
    
    waterb_risk_rivers_merged[[i]] =     dplyr::bind_rows(rivers_products_b)
    }
  }

waterb_risk_rivers_merged = waterb_risk_rivers_merged |> 
  dplyr::bind_rows()

# waterb_risk_rivers_merged = waterb_risk_rivers_to_merge |> 
#  group_by(GNIS_NA) |> 
#  #Arrange the dataframe so that the highest risk section of each river comes first.
#  arrange(GNIS_NA,desc(as.numeric(InvasionRisk))) |> 
#  #Take the first value for each of our variables and set it as the default for each river (basically, homogenize the data in a nservative way), then merge polygons.
#  summarise(across(!starts_with("geo"), first))
rm(waterb_risk_rivers_to_merge)

#Reattach rivers to the overall waterb_risk spatial object.
waterb_risk = waterb_risk |> 
 filter(!paste0(WATERSH,WATERBO,GNIS_NA) %in% 
          all_of(paste0(duplicated_rivers$WATERSH,
                        duplicated_rivers$WATERBO,
                        duplicated_rivers$GNIS_NA))) |> 
 #Add in the rivers.
 bind_rows(waterb_risk_rivers_merged)
rm(duplicated_river_names); rm(duplicated_rivers); rm(waterb_risk_rivers_merged)
```

```{r add_NR_Region_name_to_waterbodies}
# nr_regions = bcmaps::nr_regions()
nr_regions = sf::read_sf("W:/CMadsen/shared_data_sets/FLNRO_Fishing_Boundaries.shp")

waterb_risk_with_region = waterb_risk |> 
 st_join(nr_regions |> dplyr::select(REGION_N),
         join = st_intersects)

waterb_risk_with_region = waterb_risk_with_region |> 
#Some waterbodies fall on the border between 2+ FLNRO regions...
 group_by(WATERSH,WATERBO,GNIS_NA) |> 
 #For each unique grouping of watershed, waterbody id and name...
 mutate(REGION_N = paste0(REGION_N, collapse = ", ")) |> 
 distinct() |> 
 ungroup()

waterb_risk = waterb_risk_with_region
rm(waterb_risk_with_region)
```

```{r drop_dry_storage_and_pac_ocean}
# Drop Dry Storage and Pacific Ocean.
waterb_risk = waterb_risk |> 
 dplyr::filter(!str_detect(GNIS_NA, '(Dry Storage|Pacific Ocean)'))
```

```{r read_in_past_year_sampling}

sample_protocol_2021 = read_excel(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/2021/Updates to protocol_waterbody ranking/2021 priority waterbodies for sampling.xlsx"),sheet = "2021 FINAL LIST")

sample_protocol_2022 = get_wb_list_from_protocol_doc(
  officer::read_docx('J:/2 SCIENCE - Invasives/SPECIES/Zebra_Quagga_Mussel/Operations/Lake Monitoring/2022/Sampling protocol/2022 Invasive Mussel Field Protocol_Draft_V1.docx')
)

sample_protocol_2023 = get_wb_list_from_protocol_doc(
  officer::read_docx('J:/2 SCIENCE - Invasives/SPECIES/Zebra_Quagga_Mussel/Operations/Lake Monitoring/2023/Sampling protocol/2023 Invasive Mussel Field Protocol_Draft_V1.docx')
)

sample_protocol_2024 = get_wb_list_from_protocol_doc(
  officer::read_docx('J:/2 SCIENCE - Invasives/SPECIES/Zebra_Quagga_Mussel/Operations/Lake Monitoring/2024/Sampling protocol/2024 Invasive Mussel Field Protocol_Final.docx')
)

past_sample_protocols = sample_protocol_2021 |> 
 mutate(sampling_year = 2021) |> 
 bind_rows(sample_protocol_2022 |> mutate(Lat = as.numeric(Lat), Long = as.numeric(Long)) |> mutate(sampling_year = 2022)) |> 
 bind_rows(sample_protocol_2023 |> mutate(Lat = as.numeric(Lat), Long = as.numeric(Long)) |> mutate(sampling_year = 2023)) |> 
  bind_rows(sample_protocol_2024 |> mutate(Lat = as.numeric(Lat), Long = as.numeric(Long)) |> mutate(sampling_year = 2024))

#We have some waterbodies with various sampling stations proposed. We just need 1 row per wb.
sample_protocol = past_sample_protocols |>
  group_by(Region,Waterbody,sampling_year) |>
  slice(1) |> 
  dplyr::ungroup()

# Some names are spelled differently between this excel sheet and the BCG Warehouse waterbody layers. Fix these manually.
sample_protocol = sample_protocol |>
  mutate(Waterbody = case_when(
    Waterbody == "Atlin lake" ~ "Atlin Lake",
    Waterbody == "Columbia River (Lower)" ~ "Columbia River",
    Waterbody == "Kinaskan" ~ "Kinaskan Lake",
    Waterbody == "Koocanusa Lake" ~ "Lake Koocanusa",
    Waterbody == "Marquart & Lunbom Lakes" ~ "Marquart Lake",
    T ~ Waterbody
  )) |>
  group_by(Region,Waterbody,sampling_year) |>
  slice(1) |> 
  dplyr::ungroup()

# # Rename regions to match with inspection data regions.
# sample_protocol = sample_protocol |> 
#   dplyr::mutate(Region = replace(Region, Region == 'Okanagan', 'Thompson-Okanagan')) |> 
#   dplyr::mutate(Region = str_replace(Region, 'Thompson-Nicola', 'Thompson-Okanagan')) |> 
#   dplyr::mutate(Region = str_replace(Region, '(Vancouver Island|Lower Mainland)', 'South Coast')) |> 
#   dplyr::mutate(Region = str_replace(Region, 'Kootenay', 'Kootenay-Boundary'))

sample_protocol = sample_protocol |> 
  dplyr::select(Region, Waterbody, sampling_year) |> 
  dplyr::mutate(to_sample = TRUE) |> 
  dplyr::mutate(sampling_year = paste0('on_',sampling_year,'_protocol_list')) |> 
  pivot_wider(names_from = sampling_year, values_from = to_sample, values_fill = FALSE) |> 
  ungroup() |> 
  dplyr::rename(GNIS_NA = Waterbody,
                REGION_NAME = Region) |> 
  dplyr::select(REGION_NAME,GNIS_NA,paste0('on_',2021:my_opts$year,'_protocol_list'))

rm(past_sample_protocols)
```

```{r add_past_sampling_to_waterb_risk}
#Do a name match between "waterb_risk" and the protocol list
waterb_risk = waterb_risk |>
  dplyr::rename(REGION_NAME = REGION_N) |>
 #Name join
  left_join(sample_protocol, by = join_by(GNIS_NA, REGION_NAME)) |> 
  dplyr::mutate(dplyr::across(dplyr::contains("_protocol_list"),\(x) tidyr::replace_na(x, F))) |> 
  ungroup()
rm(sample_protocol); rm(sample_protocol_2021); rm(sample_protocol_2022); rm(sample_protocol_2023); rm(sample_protocol_2024)
```

```{r add_lake_monitoring_lab_results}
#2022 Lake monitoring lab results.
lakes_monitored_2022 = read_excel(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/2022/Lab Analysis/Final report and data/BC Veliger Sampling Inventory 2022_FinalReport.xlsx")) |> 
  dplyr::mutate(`Total Volume Sampled (L)` = as.numeric(`Total Volume Sampled (L)`)) |> 
  dplyr::mutate(Year = 2022)

#2023 Lake monitoring lab results.
lakes_monitored_2023 = read_excel(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/2023/Lab results/BC Veliger Sampling Inventory 2023_11-20_Final Report.xlsx")) |> 
  dplyr::filter(!is.na(Waterbody),
                !stringr::str_detect(Waterbody,'[0-9]+'),
                Waterbody != 'waterbodies') |> 
  dplyr::mutate(Waterbody = stringr::str_squish(Waterbody)) |> 
  dplyr::mutate(Waterbody = case_when(
    Waterbody == 'Arrow Lake, Lower' ~ 'Lower Arrow Lake',
    Waterbody == 'Arrow Lake, Upper' ~ 'Upper Arrow Lake',
    Waterbody == 'Enid Lake' ~ 'Lake Enid',
    Waterbody == 'Koocanusa Lake' ~ 'Lake Koocanusa',
    Waterbody == 'Kootenay River (Nelson)' ~ 'Kootenay River',
    Waterbody == 'Lower Fraser River' ~ 'Fraser River',
    Waterbody == 'Lac La Hache' ~ 'Lac la Hache',
    Waterbody == 'Lower Kettle River' ~ 'Kettle River',
    Waterbody == 'Upper Kettle River' ~ 'Kettle River',
    Waterbody == 'Norbury Lake' ~ 'Norbury Lakes',
    Waterbody == 'Winderemere Lake' ~ 'Windermere Lake',
    Waterbody == "Pend d'Oreille River" ~ "Pend-d'Oreille River",
    T ~ Waterbody
  )) |> 
  dplyr::mutate(`# of Plankton tows` = as.character(`# of Plankton tows`)) |> 
  dplyr::mutate(Year = 2023)

# 2024 
lakes_monitored_2024 = read_excel(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/2024/Lab results/Final report/BC Veliger Sampling Inventory 2024_FINAL REPORT.xlsx")) |> 
  dplyr::mutate(`Lat (Decimal degrees)` = as.numeric(`Lat (Decimal degrees)`),
                `Long (Decimal degrees)` = as.numeric(`Long (Decimal degrees)`)) |> 
  dplyr::mutate(`Water temperature (°C)*` = as.character(`Water temperature (°C)*`),
                `pH (in water column)` = as.character(`pH (in water column)`),
                `Total Volume Sampled (L)` = as.numeric(`Total Volume Sampled (L)`),
                `Sample pH (at the time preserved)` = as.character(`Sample pH (at the time preserved)`)) |> 
  dplyr::filter(!is.na(Waterbody),
                !stringr::str_detect(Waterbody,'[0-9]+'),
                Waterbody != 'waterbodies') |> 
  dplyr::mutate(Waterbody = stringr::str_squish(Waterbody)) |> 
  dplyr::mutate(Waterbody = case_when(
    Waterbody == 'Alouette' ~ 'Alouette Lake',
    Waterbody == 'Leighton' ~ 'Leighton Lake',
    Waterbody == 'Wahleach' ~ 'Wahleach Lake',
    Waterbody == 'Peter-Hope' ~ 'Peter Hope Lake',
    Waterbody %in% c('St. Marys Lake',"St. Mary's Lake") ~ 'St. Mary Lake',
    Waterbody %in% c('Surveyors Lake',"Surveyor's Lake") ~ 'Surveyors Lake',
    Waterbody == 'Arrow Lake, Lower' ~ 'Lower Arrow Lake',
    Waterbody == 'Arrow Lake, Upper' ~ 'Upper Arrow Lake',
    Waterbody == 'Enid Lake' ~ 'Lake Enid',
    Waterbody == 'Koocanusa Lake' ~ 'Lake Koocanusa',
    stringr::str_detect(Waterbody, 'Kootenay River \\(*') ~ 'Kootenay River',
    Waterbody == 'Lower Fraser River' ~ 'Fraser River',
    Waterbody == 'Lac La Hache' ~ 'Lac la Hache',
    Waterbody == 'Lac le Jeune' ~ 'Lac Le Jeune',
    Waterbody == 'Lower Kettle River' ~ 'Kettle River',
    Waterbody == 'Upper Kettle River' ~ 'Kettle River',
    Waterbody == 'Norbury Lake' ~ 'Norbury Lakes',
    Waterbody == 'Winderemere Lake' ~ 'Windermere Lake',
    Waterbody == "Pend d'Oreille River" ~ "Pend-d'Oreille River",
    T ~ Waterbody
  )) |> 
  dplyr::mutate(`# of Plankton tows` = as.character(`# of Plankton tows`)) |> 
  dplyr::mutate(Year = 2024)

# Find the natural resource region, add in for each wb.
lakes_monitored = lakes_monitored_2022 |> 
  dplyr::bind_rows(lakes_monitored_2023) |> 
  dplyr::bind_rows(lakes_monitored_2024) |> 
  dplyr::filter(!is.na(`Lat (Decimal degrees)`)) |> 
  st_as_sf(coords = c('Long (Decimal degrees)',
                      'Lat (Decimal degrees)'),
           crs = 4326) |>
  st_transform(3005) |> 
  st_join(nr_regions |> 
            dplyr::select(REGION_N)) |> 
            # dplyr::mutate(REGION_N = stringr::str_remove_all(REGION_N,' Natural Resource Region')) |> 
            # dplyr::mutate(REGION_N = replace(REGION_N, REGION_N == 'Northeast','Peace'))) |> 
  st_transform(4326) |> 
  dplyr::rename(GNIS_NA = Waterbody) |> 
  st_drop_geometry() |> 
  dplyr::select(GNIS_NA, REGION_N, Year) |> 
  dplyr::distinct() |> 
  dplyr::mutate(sampled = TRUE) |> 
  dplyr::mutate(Year = paste0('Sampled_in_',Year)) |> 
  pivot_wider(names_from = Year, values_from = sampled, values_fill = FALSE) |> 
  dplyr::rename(REGION_NAME = REGION_N)

waterb_risk = waterb_risk |>
  dplyr::left_join(lakes_monitored, by = join_by(GNIS_NA, REGION_NAME)) |> 
  dplyr::mutate(dplyr::across(dplyr::starts_with("Sampled_in_"),\(x) tidyr::replace_na(x, FALSE)))
rm(lakes_monitored); rm(lakes_monitored_2022); rm(lakes_monitored_2023)
```

Prior to filtering out any waterbodies, there are `r nrow(waterb_risk)` waterbodies for which we have at least one risk variable. Note that at all filtering steps, all waterbodies with at least one record of a mussel-fouled boat were retained.

Filtering Steps:

```{r full_wb_table_for_exploring_wb_values}
waterb_full_tbl = waterb_risk |> 
  sf::st_drop_geometry() |> 
  dplyr::left_join(sf::st_drop_geometry(waterb_og_values_and_bins))
```

```{r highlight_certain_waterbodies_to_keep}
waterb_risk = waterb_risk |> 
  dplyr::mutate(keep_me_always = dplyr::case_when(
    GNIS_NA %in% c('Kootenay River') ~ T,
    T ~ F
  ))
```

```{r filter_out_risk_bin_1_or_lvl_2_and_low_calcium}
#Filtering steps outlined by Cass and Martina. First batch!
waterb_risk_complete = waterb_risk

waterb_risk = waterb_risk |> 
  mutate(keep.me = case_when(
    keep_me_always ~ 'keep',
    paste0(GNIS_NA,WATERSH) %in% paste0(waterb_with_any_mf$GNIS_NA,waterb_with_any_mf$WATERSH) ~ 'keep',
    InvasionRisk >= 2 ~ "keep",
    # GNIS_NA == 'Jewel Lake' ~ 'keep',
    InvasionRisk == 1 ~ "drop",
    # InvasionRisk == 2 & calcium_bin == 0 & calcium_data == "Data present" ~ "drop",
    T ~ "keep"
  )) |> 
  filter(keep.me == "keep") |> 
  dplyr::select(-keep.me)
```

1. Remove waterbodies with an estimated risk of ZQ Mussel invasion in bin 1 (`r nrow(waterb_risk)` waterbodies retained).

```{r remove_use_below_1.5}
#And now the actual second filtering step!
waterb_risk = waterb_risk |>
 filter(Use > 1.5 | GNIS_NA %in% unique(waterb_with_any_mf$GNIS_NA) & WATERSH %in% unique(waterb_with_any_mf$WATERSH) | keep_me_always)
```

2. Remove waterbodies with a binned value of the 'Use' category up to 1.5 (`r nrow(waterb_risk)` waterbodies retained)

```{r remove_if_use_under_2_unless_highrisk_inspection}
# waterb_risk = waterb_risk |> 
#   filter(Highrisk > 0 | Use > 2 | GNIS_NA %in% unique(waterb_with_any_mf$GNIS_NA) & WATERSH %in% unique(waterb_with_any_mf$WATERSH))
# 3. Remove waterbodies with a binned value of the 'Use' category up to 2 unless they have at least one high-risk inspection (`r nrow(waterb_risk)` waterbodies retained)
```


```{r remove_kenai_creek_and_merge_Kinbasket_Lake}
#Third filtering step: remove Kenai Creek and check for duplicated lakes.
waterb_risk = waterb_risk |> 
 filter(GNIS_NA != "Kenai Creek")

#Just for Kinbasket Lake - merge the two rows, taking higher risk value data for most fields.
waterb_risk = waterb_risk |> 
 filter(GNIS_NA != "Kinbasket Lake") |> 
 bind_rows(
 waterb_risk |> 
 filter(GNIS_NA == "Kinbasket Lake") |> 
 group_by(GNIS_NA) |> 
 arrange(GNIS_NA,desc(as.numeric(InvasionRisk))) |> 
 summarise(across(!starts_with("geo"), first)) |> 
 mutate(REGION_G = "4, 7O", REGION_N = "Kootenay, Omineca")
 )

# I seem to have duplicate columns from the FLNRORD natural resource regions...
# drop those here.
waterb_risk = waterb_risk |> 
 dplyr::select(-REGION_N, -REGION_G)
```

3. Remove Kenai Creek and merge multiple rows for Kinbasket Lake into one row (`r nrow(waterb_risk)` waterbodies retained)

```{r filter_away_calcium_level_no_risk}
waterb_risk = waterb_risk |> 
  dplyr::filter(calcium_bin > 0 | keep_me_always)
```

4. Remove waterbodies with a calcium risk bin of 0 (which equates to "No Risk"); (`r nrow(waterb_risk)` waterbodies retained)

```{r write_to_W_drive}
#Write to file
waterb_risk = waterb_risk |>
          ungroup() |>
          # dplyr::select(-name_risk_ranking) |>
          rename(OnProtocolList2021 = on_2021_protocol_list,
                 OnProtocolList2022 = on_2022_protocol_list,
                 OnProtocolList2023 = on_2023_protocol_list,
                 OnProtocolList2024 = on_2024_protocol_list) |>
          mutate(InvasionRisk = as.numeric(InvasionRisk),
                 EVRisk = as.numeric(EVRisk)#,
                 # OnProtocolList2022 = as.character(OnProtocolList2022),
                 # Sampled2022 = as.character(Sampled2022)
                 )

if(update_remote_output_files){
  if(!dir.exists(paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/"))){
    dir.create(paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/"))
  }
  write_sf(waterb_risk,
           paste0(my_opts$remote_spatial_data,"Projects/ZQMussels/",my.year," IMDP Final Report/data/spatial/waterbodies_zqm_shortlist_risk_estimates.gpkg"))
}
```

```{r make_final_table}
# Add in waterbody surface area
waterb_risk = waterb_risk |> 
  mutate(wb_area_km_2 = as.numeric(sf::st_area(geom))/1000000)

# Find centroid point of each waterbody. Add to table.
waterbody_risk_table = st_centroid(waterb_risk) |> 
 ungroup() |> 
 st_transform(crs = 4326) |> 
 mutate(Lat = st_coordinates(geom)[,2],
        Long = st_coordinates(geom)[,1]) |>
 st_drop_geometry() |> 
 arrange(desc(InvasionRisk))
```

```{r add_unbinned_versions_of_variables_to_waterb_risk_table}
waterbody_risk_table = waterbody_risk_table |> 
  # dplyr::mutate(EVRisk = as.character(EVRisk)) |> 
  left_join(sf::st_drop_geometry(waterb_og_values_and_bins))
```

```{r calculate_ratio_of_total_inspections_to_angler_days}
waterbody_risk_table = waterbody_risk_table |> 
  dplyr::mutate(inspections_angler_days_ratio = TotalInspections/days_fished_2023_survey)
```

```{r}
#Replace the unique ID numbers of subwatersheds with their names.
waterbodies = read_sf(paste0(my_opts$remote_spatial_data,'shared_data_sets/WatershedGroups_lowres.shp')) |> 
 st_drop_geometry() |> 
 reframe(WATERSH = WATERSHED_, WatershedName = WATERSHE_1)

waterbody_risk_table = waterbody_risk_table |> 
 left_join(waterbodies) |> 
 dplyr::select(WatershedName, everything()) |> 
 dplyr::select(-WATERSH)
  
```

Here is the shortlist (which has `r nrow(waterbody_risk_table)` rows.)

```{r show_risk_table_as_DT}
DT::datatable(
  waterbody_risk_table
)
```

And here is a histogram of the waterbody surface area values.

```{r wb_surface_area_histogram}
ggplot(waterbody_risk_table) +
  geom_histogram(aes(wb_area_km_2)) +
  labs(x = "Waterbody Surface Area (km^2)", y = "Number of Waterbodies")
```

```{r write_excel_files_to_disk}
# ---------------- #
#      NOTE        #
# ---------------- #
# To run the code below (which generates an excel file wherein high-risk lakes that have not yet been sampled are highlighted in yellow), it will be necessary to first un-comment the section above to see which lakes were sampled in the previous year, as this section adds 2 columns to the waterbodies_risk table which are necessary for the code below to run successfully.

my.wb = createWorkbook("waterbodies_risk")

addWorksheet(wb = my.wb, sheetName = "Risk Estimates")

writeDataTable(wb = my.wb,
               sheet = "Risk Estimates",
               x = waterbody_risk_table,
               withFilter = T)

# #Highlight any high-risk waterbodies that have not been sampled.
# posStyle <- createStyle(fontColour = "#fa3232", bgFill = "#ffff31")
# conditionalFormatting(wb=my.wb, sheet='Risk Estimates', cols=which(names(waterbody_risk_table) == paste0('OnProtocolList',my_opts$year)),
#                      rows=1:nrow(waterbody_risk_table), rule="== FALSE",
#                      type = "expression",style = posStyle)
# conditionalFormatting(wb=my.wb, sheet='Risk Estimates', cols=which(names(waterbody_risk_table) == paste0('Sampled',my_opts$year)),
#                      rows=1:nrow(waterbody_risk_table), rule="== FALSE",
#                      type = "expression",style = posStyle)

#Automatically adjust cell widths based on longest cell contents OR header.
width_vec <- apply(waterbody_risk_table, 2,
                  function(x) max(nchar(as.character(x)) + 2, na.rm = TRUE))
width_vec_header <- nchar(colnames(waterbody_risk_table)) + 2
max_vec_header <- pmax(width_vec, width_vec_header)
setColWidths(my.wb, "Risk Estimates",
            cols = 1:ncol(waterbody_risk_table), widths = max_vec_header)

#Write a second sheet for metadata.
addWorksheet(wb = my.wb,
            sheetName = "Metadata")

explanation_col = c("Name of waterbody's subwatershed according to BCG warehouse subwatershed layer",
                                             "Unique identifying number of waterbody polygon in BCG Warehouse lake, river or man-made waterbody layers",
                                             "Name of waterbody in BCG Warehouse lake, river or man-made waterbody layers",
                                             "Risk estimate for Use variables (total inspections, angler days, and number of marinas)",
                                             "Risk estimate for High-risk variables (High-risk inspections minus Transport Canada operation restrictions)",
                                             "The number of high-risk inspections",
                                             "Risk estimate (1 or 0) depending on mussel-fouled inspections or no",
                                             "The summed dam capacity at the waterbody scale",
                                             "The number of distinct species-at-risk in the waterbody",
                                             "Subwatershed-scale dissolved calcium levels (<8 is low risk, 8-20 medium risk, 20+ high risk); bins are 0, 1 and 2",
                                             "If calcium bin is 0, is this due to the calcium level being <8, or because that subwatershed lacked calcium data?",
                                             "Overall risk estimate calculated by combining the use, high-risk, mussel-fouled and calcium_bin risk estimates",
                                             "As last column, but binned into 3 bins",
                                             "FLRNO region code",
                                             "Was the waterbody included in the 2021 list of priority lakes to sample?",
                                             "Was the waterbody included in the 2022 list of priority lakes to sample?",
                                             "Was the lake sampled in 2022?",
                                             "Was the lake sampled in 2023?",
                                             "Latitude of centroid of polygon",
                                             "Longitude of centroid of polygon")

explanation_col = c(explanation_col, 
                    rep('Unbinned variable', 
                        ncol(waterbody_risk_table) - length(explanation_col))
)

writeDataTable(wb = my.wb, sheet = "Metadata",
              x = data.frame(Variable = names(waterbody_risk_table),
                             Explanation = explanation_col),
              withFilter = F)

setColWidths(my.wb, "Metadata",
            cols = 1, widths = 17)

# saveWorkbook(my.wb,
#              "output/waterbodies with risk estimates post filtering.xlsx",
#              overwrite = T)

if(!dir.exists(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/"))) {
  dir.create(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/"), recursive = T)
}

saveWorkbook(my.wb,
            paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/waterbodies with risk estimates post filtering_",my.year,".xlsx"),
            overwrite = T)
```

```{r final_waterbody_list}

# Please see note in above chunk. It applies to this section as well, I'm fairly sure.

final_waterbody_list = waterbody_risk_table |>
 ungroup() |>
  rename(Region = REGION_NAME,
         Waterbody = GNIS_NA,
         InvasionRisk_bin = InvasionRisk,
    EVRisk_bin = EVRisk) |> 
  dplyr::mutate(
    `Veliger Sampling Frequency` = "Undecided",
    Lat = round(Lat, 4),
    Long = round(Long, 4)
  ) |> 
  dplyr::mutate(Region = stringr::str_remove_all(Region, ' Natural Resource Region'))

fwl_rivers = final_waterbody_list |>
 filter(str_detect(Waterbody, "River"))

fwl_lakes = final_waterbody_list |>
 filter(!str_detect(Waterbody, "River"))

#Change coords for rivers to NA (delete fields!)
fwl_rivers$Lat = NULL
fwl_rivers$Long = NULL

final_waterbody_list = fwl_lakes |>
 arrange(Waterbody) |>
 bind_rows(fwl_rivers |> arrange(Waterbody)) |>
 arrange(Waterbody)

# Add in the sampling frequency!
final_waterbody_list = final_waterbody_list |> 
  dplyr::mutate(Sampling_Frequency = dplyr::case_when(
    InvasionRisk_bin == 3 | Highrisk == 3 ~ "Plankton Bi-weekly", 
    Use < 2 ~ "Substrate Only",
    T ~ "Plankton Monthly"))

final_waterbody_list = final_waterbody_list |> 
  dplyr::arrange(Sampling_Frequency, InvasionRisk_bin, Use)

openxlsx::write.xlsx(final_waterbody_list,
                    paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/Final waterbody list based on risk estimates_",my.year,".xlsx"),
                    overwrite = T)

# Add a sheet to this doc that has only the variables needed for the table in word.
final_wb_wb = openxlsx::loadWorkbook(file = paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/Final waterbody list based on risk estimates_",my.year,".xlsx"))

openxlsx::addWorksheet(final_wb_wb, sheetName = "table_for_word")

openxlsx::writeDataTable(final_wb_wb,
                         "table_for_word", 
                         x = final_waterbody_list |>
                           dplyr::select(Region,Waterbody,Sampling_Frequency,Lat,Long)
)

openxlsx::saveWorkbook(final_wb_wb, 
                       file = paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/Final waterbody list based on risk estimates_",my.year,".xlsx"),
                       overwrite = T)
```

```{r little_exploration_bit}
# How many wbs are in bin 1 for inspections and bin 3 for angler days?
insp_angler_binned_tbl = final_waterbody_list |> 
  dplyr::select(Region,Waterbody,TotalInspections,days_fished_2023_survey) |> 
  pivot_longer(-c(Region,Waterbody), names_to = "Variable") |> 
  dplyr::left_join(
    naturalbreaks
  ) |> 
  dplyr::mutate(
    variable_bin = dplyr::case_when(
      value > Natural_Break_1 & value <= Natural_Break_2 ~ 1,
      value > Natural_Break_2 & value <= Natural_Break_3 ~ 2,
      value > Natural_Break_3 & value <= Natural_Break_4 ~ 3,
      is.na(value) ~ NA,
      T ~ 99
    )) |> 
  dplyr::select(Region,Waterbody,Variable,variable_bin) |> 
  pivot_wider(names_from = Variable, values_from = variable_bin) 

insp_angler_binned_tbl |>
  dplyr::count(TotalInspections,days_fished_2023_survey, sort = T) |> 
  knitr::kable()

# insp_angler_binned_tbl |> 
#   dplyr::filter(TotalInspections == 1, days_fished_2023_survey == 3)

```

```{r make_comparison_files}
# last_year_from_excel = read_excel(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/2024/Prioritization model/Final waterbody list based on risk estimates_2024.xlsx"))

final_waterbody_list = final_waterbody_list |> 
  mutate(region_wb = paste0(Region,'-',Waterbody))

last_year = get_wb_list_from_protocol_doc(
  officer::read_docx('J:/2 SCIENCE - Invasives/SPECIES/Zebra_Quagga_Mussel/Operations/Lake Monitoring/2024/Sampling protocol/2024 Invasive Mussel Field Protocol_Final.docx')
) |> 
  mutate(Waterbody = stringr::str_replace(Waterbody,"’","'")) |> 
  mutate(region_wb = paste0(Region,'-',Waterbody))
  
last_year$only_on_last_years_list = FALSE

for(i in 1:nrow(last_year)){
  relevant_wbs = final_waterbody_list[final_waterbody_list$Waterbody == last_year[i,]$Waterbody,]
  
  if(nrow(relevant_wbs) == 0) {
    last_year[i,]$only_on_last_years_list = TRUE
  } else {
    if(str_detect(paste0(relevant_wbs$Region,collapse=''),last_year[i,]$Region)){
      last_year[i,]$only_on_last_years_list = FALSE
    }
  }
}

final_waterbody_list$new_on_this_years_list = FALSE

for(i in 1:nrow(final_waterbody_list)){
  
  relevant_wbs = last_year[last_year$Waterbody == final_waterbody_list[i,]$Waterbody,]
  
  if(nrow(relevant_wbs) == 0) {
    final_waterbody_list[i,]$new_on_this_years_list = TRUE
  } else {
    if(str_detect(paste0(relevant_wbs$Region,collapse=''),final_waterbody_list[i,]$Region)){
      final_waterbody_list[i,]$new_on_this_years_list = FALSE
    }
  }
}

waterbodies_dropped_from_last_year = last_year[last_year$only_on_last_years_list == T,]
waterbodies_added_this_year = final_waterbody_list[final_waterbody_list$new_on_this_years_list == T,]

# Add a piece describing which of the lakes ONLY in last year's protocol (and dropped from this year's protocol)
# were actually sampled.
waterbodies_dropped_from_last_year$Sampled_in_2024 = "FALSE"
for(i in 1:nrow(waterbodies_dropped_from_last_year)){
  relevant_wbs = waterb_full_tbl[waterb_full_tbl$GNIS_NA == waterbodies_dropped_from_last_year[i,]$Waterbody,]
  if(nrow(relevant_wbs) != 0) {
    if(nrow(relevant_wbs) == 1){
      waterbodies_dropped_from_last_year[i,]$Sampled_in_2024 = as.character(relevant_wbs$Sampled_in_2024)
    } else {

      point_for_search = waterbodies_dropped_from_last_year[i,] |> 
        sf::st_as_sf(coords = c("Long","Lat"), crs = 4326)
      
      watershed_groups = bcdc_query_geodata('freshwater-atlas-watershed-groups') |> 
        filter(WATERSHED_GROUP_ID %in% relevant_wbs$WATERSH) |> 
        collect() |> 
        sf::st_transform(4326)
      
      watershed_group_id_filter = watershed_groups |> 
        sf::st_filter(point_for_search) |> 
        pull(WATERSHED_GROUP_ID)
      
      if(length(watershed_group_id_filter) == 0){
        waterbodies_dropped_from_last_year[i,]$Sampled_in_2024 = "Unknown"
      } else {
      waterbodies_dropped_from_last_year[i,]$Sampled_in_2024 = as.character(relevant_wbs[relevant_wbs$WATERSH ==  watershed_group_id_filter,]$Sampled_in_2024)
      }
    }
  }
}

# Save results.
my.wb = createWorkbook("annual_sampling_lists_comparison")
addWorksheet(wb = my.wb, sheetName = "Only in last year")
addWorksheet(wb = my.wb, sheetName = "Only in this year")

writeDataTable(wb = my.wb,
                   sheet = "Only in last year",
                   x = waterbodies_dropped_from_last_year,
                   withFilter = T)
writeDataTable(wb = my.wb,
                   sheet = "Only in this year",
                   x = waterbodies_added_this_year |> arrange(Waterbody),
                   withFilter = T)
saveWorkbook(my.wb,
            paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/Sampling_list_comparison_",my.year,"_with_",my.year-1,".xlsx"),
              overwrite = T)
```

This year, there are `r nrow(final_waterbody_list)` waterbodies on our protocol list. Last year, there were `r nrow(last_year)`.

To identify the waterbodies for bi-weekly sampling, the final waterbody list (`r nrow(final_waterbody_list)`) was sorted by the overall invasion risk and also by the high-risk category score. All waterbodies with a score of 3 for invasion risk (the highest possible risk ranking) and/or a score of 3 for the “high risk” category (the highest possible risk ranking) were selected for bi-weekly sampling. This resulted in `r nrow(waterbody_risk_table |> filter(InvasionRisk == 3 | Highrisk == 3))` waterbodies being selected; one additional waterbody, the Kootenay River, was also selected because it flows into Idaho (of special concern following a reporting of Zebra-Quagga Mussels) and then back into BC. This resulted in a total of `r nrow(waterbody_risk_table |> filter(InvasionRisk == 3 | Highrisk == 3))+1` waterbodies in total being selected for bi-weekly sampling.

