---
title: "2023 Metabase Data Exploration"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
---
```{r prep, include = F}
library(tidyverse)
library(plotly)
library(ggrepel)

knitr::opts_chunk$set(echo=F,warning=F,message = F)

# Read in a year of metabase data.
# setwd("C:/Users/CMADSEN/Downloads/LocalR/ZQMussels/")

dat = read_csv('J:/2 SCIENCE - Invasives/SPECIES/Zebra_Quagga_Mussel/Operations/Watercraft Inspection Data/Raw inspection data for sharing (all years)/Clean files all years/metabase_2023.csv')

```

## Methodology for this assessment 

Destination information is gathered from the following fields, in order of preference (e.g. if there is a waterbody name and closest city, the waterbody name is chosen):

- Destination Waterbody 1 Name
- Destination Waterbody 1 Closest City
- Destination Major City
- Destination Waterbody Other Info

Source information is gathered in a similar way, from the corresponding fields:

- Previous Waterbody 1 Name
- Previous Waterbody 1 Closest city
- Previous Major City
- Previous Waterbody Other Info

## Figures {.tabset}

### Destination Waterbody Info Completeness by Station

```{r}
# Make some plots!

## Where are boats coming from and where are they heading to?

# Missing information for destination waterbody name / closest city.
dest_dat = dat |> 
  filter(!str_detect(Station,"Scheduled")) |> 
  mutate(`Destination Major City` = replace(`Destination Major City`,`Destination Major City` == 'None', NA)) |> 
  mutate(destination_any_info = coalesce(`Destination Waterbody 1 Name`,`Destination Waterbody 1 Closest City`,`Destination Major City`,`Destination Waterbody 1 Other Details`)) |> 
  count(Station, destination_any_info, sort = T) |> 
  mutate(something_known = !is.na(destination_any_info)) |> 
  group_by(Station, something_known) |> 
  mutate(total_n = sum(n)) |> 
  dplyr::select(something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  group_by(Station) |> 
  mutate(total_all = sum(total_n)) |> 
  ungroup() |> 
  arrange(desc(total_all)) |> 
  mutate(Station = forcats::fct_inorder(Station))

plot_ly(dest_dat,
        colors = c('orange','lightblue')) |> 
  add_bars(x = ~Station, y = ~total_n, color = ~something_known) |> 
  layout(yaxis = list(title = 'Number of Inspections'),
         xaxis = list(title = 'Station'),
         legend = list(title=list(text='Any Destination Info')),
         barmode = 'stack',
         title = "Destination Data Field Completeness by Station"
  )
         
         
         
# p = ggplot(dest_dat) + 
#            geom_col(aes(x = something_known, y = total_n, fill = `High Risk AIS Ind`)) + 
#            labs(y = 'Number of Records', title = 'Data for Destination WB name / closest city / "destination closest city"', x = '')
#          
# ggplotly(p)
```

### Previous Waterbody Info Completeness by Station

```{r}
# Source information
# Missing information for source waterbody name / closest city.
prev_dat = dat |> 
    filter(!str_detect(Station,"Scheduled")) |> 
  mutate(`Previous Major City` = replace(`Previous Major City`,`Previous Major City` == 'None', NA)) |> 
  mutate(Previous_any_info = coalesce(`Previous Waterbody 1 Name`,`Previous Waterbody 1 Closest City`,`Previous Major City`,`Previous Waterbody 1 Other Details`)) |> 
count(Station, Previous_any_info, sort = T) |> 
  mutate(something_known = !is.na(Previous_any_info)) |> 
  group_by(Station,something_known) |> 
  mutate(total_n = sum(n)) |> 
  dplyr::select(Station,something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  group_by(Station) |> 
  mutate(total_all = sum(total_n)) |> 
  ungroup() |> 
  arrange(desc(total_all)) |> 
  mutate(Station = forcats::fct_inorder(Station))

plot_ly(prev_dat,
        colors = c('orange','lightblue')) |> 
  add_bars(x = ~Station, y = ~total_n, color = ~something_known) |> 
  layout(yaxis = list(title = 'Number of Inspections'),
         xaxis = list(title = 'Station'),
         legend = list(title=list(text='Any Source Info')),
         barmode = 'stack',
         title = "Source Location Data Field Completeness by Station"
  )
# p = ggplot(prev_dat) + 
#   geom_col(aes(x = something_known, y = total_n, fill = `High Risk AIS Ind`)) + 
#   labs(y = 'Number of Records', title = 'Data for either Previous WB name / closest city / "previous major city"', x = '')
# 
# ggplotly(p)
```

### Destination Waterbody Info Completeness by Inspector

```{r}
dest_dat = dat |> 
  mutate(Email = stringr::str_remove_all(Email, '@gov.*')) |> 
    filter(!str_detect(Station,"Scheduled")) |> 
  mutate(`Destination Major City` = replace(`Destination Major City`,`Destination Major City` == 'None', NA)) |> 
  mutate(destination_any_info = coalesce(`Destination Waterbody 1 Name`,`Destination Waterbody 1 Closest City`,`Destination Major City`,`Destination Waterbody 1 Other Details`)) |> 
  count(Email, destination_any_info, sort = T) |> 
  mutate(something_known = !is.na(destination_any_info)) |> 
  group_by(Email, something_known) |> 
  mutate(total_n = sum(n,na.rm=T)) |> 
  dplyr::select(Email, something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  arrange(Email) |> 
  mutate(something_known = ifelse(something_known,'something_known','nothing_known')) |> 
  pivot_wider(names_from = something_known, values_from = total_n) |> 
  mutate(nothing_known = replace_na(nothing_known, 0)) |> 
  mutate(data_cov_perc = round(100 * (something_known / (nothing_known + something_known)),1))
  
# plot_ly(dest_dat) |> 
#   add_bars(y = ~reorder(Email, data_cov_perc), x = ~data_cov_perc) |> 
#   layout(yaxis = list(title = 'Data Coverage (% of records with dest. info)'),
#          xaxis = list(title = 'Inspector Email'),
#          legend = list(title=list(text='Any Source Info')),
#          barmode = 'stack',
#          title = 'Data Coverage of Destination Data Fields'
#   )
ggplot(dest_dat) + 
  geom_col(aes(y = reorder(Email, data_cov_perc),
               x = data_cov_perc)) + 
  labs(y = 'Data Coverage (% of records with destination info)',
       x = 'Inspector Email',
       title = 'Data Coverage of Destination Data Fields') + 
  scale_x_continuous(breaks = seq(0,100,10))
```

### Previous Waterbody Info Completeness by Inspector

```{r}
prev_dat = dat |> 
  mutate(Email = stringr::str_remove_all(Email, '@gov.*')) |> 
    filter(!str_detect(Station,"Scheduled")) |> 
  mutate(`Previous Major City` = replace(`Previous Major City`,`Previous Major City` == 'None', NA)) |> 
  mutate(Previous_any_info = coalesce(`Previous Waterbody 1 Name`,`Previous Waterbody 1 Closest City`,`Previous Major City`,`Previous Waterbody 1 Other Details`)) |> 
count(Email, Previous_any_info, sort = T) |> 
  mutate(something_known = !is.na(Previous_any_info)) |> 
  group_by(Email,something_known) |> 
  mutate(total_n = sum(n)) |> 
  dplyr::select(Email, something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  arrange(Email) |> 
  mutate(something_known = ifelse(something_known,'something_known','nothing_known')) |> 
  pivot_wider(names_from = something_known, values_from = total_n) |> 
mutate(data_cov_perc = round(100 * (something_known / (nothing_known + something_known)),1))
  
ggplot(prev_dat) + 
  geom_col(aes(y = reorder(Email, data_cov_perc),
               x = data_cov_perc)) + 
  labs(y = 'Data Coverage (% of records with source info)',
       x = 'Inspector Email',
       title = 'Data Coverage of Source Data Fields') + 
  scale_x_continuous(breaks = seq(0,100,10))

# plot_ly(prev_dat) |> 
#   add_bars(y = ~reorder(Email,data_cov_perc), x = ~data_cov_perc) |> 
#   layout(yaxis = list(title = 'Data Coverage (% of records with source info)'),
#          xaxis = list(title = 'Inspector Email'),
#          legend = list(title=list(text='Any Source Info')),
#          barmode = 'stack',
#          title = 'Data Coverage of Source Data Fields'
#   )
```

### Time of Inspection

```{r}
toi_dat = dat |> 
  filter(!str_detect(Station,"Scheduled")) |>
  mutate(something_known = !is.na(`Inspection Time`)) |> 
  group_by(Station, something_known) |> 
  reframe(total_n = sum(something_known,na.rm=T)) |> 
  # dplyr::select(Station, something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  arrange(Station) |> 
  mutate(something_known = ifelse(something_known,'something_known','nothing_known')) |> 
      pivot_wider(names_from = something_known, values_from = total_n)
  
if(length(ncol(toi_dat)) > 2){
  toi_dat = toi_dat |> 
    mutate(nothing_known = replace_na(nothing_known, 0))
}

  toi_dat |> 
    pivot_longer(-Station) |> 
    group_by(Station) |> 
    mutate(total_records = sum(value)) |> 
    mutate(proportion = round(100* (value / total_records) )) |> 
    ggplot() + 
    geom_col(aes(y = proportion,
                 x = Station,
                 fill = name)) + 
    scale_fill_manual(values = c('nothing_known' = 'orange','something_known' = 'lightblue')) + 
    labs(y = 'Data Coverage (%)',
         x = 'Station',
         fill = 'Some Info',
         title = 'Data Coverage of Inspection Time') + 
    theme(axis.text.x =  element_text(angle = 45, hjust = 1, vjust = 1))
```

### Previous Inspection TRUE / FALSE

```{r}
prev_insp_dat = dat |> 
  filter(!str_detect(Station,"Scheduled")) |>
  dplyr::rename(previous_inspection_ind = `Previous Inspection Ind`) |> 
  count(Station, previous_inspection_ind, sort = T) |> 
  mutate(something_known = !is.na(previous_inspection_ind)) |> 
  group_by(Station, something_known) |> 
  mutate(total_n = sum(n,na.rm=T)) |> 
  dplyr::select(Station, something_known, total_n) |> 
  distinct() |> 
  ungroup() |> 
  arrange(Station) |> 
  mutate(something_known = ifelse(something_known,'something_known','nothing_known')) |> 
      pivot_wider(names_from = something_known, values_from = total_n)
  
if(length(ncol(prev_insp_dat)) > 2){
  prev_insp_dat = prev_insp_dat |> 
    mutate(nothing_known = replace_na(nothing_known, 0))
}

  prev_insp_dat |> 
    pivot_longer(-Station) |> 
    group_by(Station) |> 
    mutate(total_records = sum(value)) |> 
    mutate(proportion = round(100* (value / total_records) )) |> 
    ggplot() + 
    geom_col(aes(y = proportion,
                 x = Station,
                 fill = name)) + 
    scale_fill_manual(values = c('nothing_known' = 'orange','something_known' = 'lightblue')) + 
    labs(y = 'Data Coverage (%)',
         x = 'Station',
         fill = 'Some Info',
         title = 'Data Coverage of Previous Inspection Ind') + 
    theme(axis.text.x =  element_text(angle = 45, hjust = 1, vjust = 1))
```

### Waterbody Complexity

```{r}
wc_dat = dat |> 
    filter(!str_detect(Station,"Scheduled")) |>
  dplyr::select(Station, `Watercraft Risk Assessment ID`, `Non Motorized Counter`, 
                `Simple Counter`, `Complex Counter`,
                `Very Complex Counter`) |> 
  pivot_longer(-c(Station,`Watercraft Risk Assessment ID`)) |> 
  group_by(Station,`Watercraft Risk Assessment ID`) |> 
  mutate(counter_sum = sum(value)) |> 
  ungroup() |> 
  mutate(data_present = ifelse(counter_sum == 0, 'absent', 'present')) |> 
  count(Station, data_present) |> 
  group_by(Station) |> 
  mutate(total_n = sum(n)) |> 
  ungroup() |> 
  arrange(desc(total_n)) |> 
  mutate(Station = as.factor(Station)) |> 
  mutate(Station = forcats::fct_inorder(Station))

plot_ly(wc_dat,
        colors = c('orange','lightblue')) |> 
  add_bars(x = ~Station, y = ~n, color = ~data_present) |> 
  layout(yaxis = list(title = 'Number of Inspections'),
         xaxis = list(title = 'Station'),
         legend = list(title=list(text='Any Complexity Info')),
         barmode = 'stack',
         title = "Watercraft Complexity Field Completeness by Station"
  )
```

### Watercraft Complexity Info Completeness by Station

```{r}
# Do we have at least one number in the watercraft type columns?
wc_type_dat = dat |> 
    filter(!str_detect(Station,"Scheduled")) |>
  dplyr::select(Station,Email,`Non Motorized Counter`,`Simple Counter`,`Complex Counter`,`Very Complex Counter`) |> 
  mutate(at_least_something = `Non Motorized Counter` + `Simple Counter` + `Complex Counter` + `Very Complex Counter` > 0) |> 
  count(Station,at_least_something) |> 
  group_by(Station) |> 
  mutate(total_inspections = sum(n)) |> 
  mutate(data_cov_perc = 100 * n / (total_inspections)) |> 
  ungroup() |> 
  arrange(desc(data_cov_perc)) |> 
  mutate(Station = forcats::fct_inorder(Station))

plot_ly(wc_type_dat,
        colors = c('orange','lightblue')
        ) |> 
  add_bars(x = ~Station, y = ~data_cov_perc, color = ~at_least_something) |> 
  layout(yaxis = list(title = 'Data Coverage (% watercrafts with type info)'),
         xaxis = list(title = 'Station'),
         legend = list(title=list(text='Any Info')),
         barmode = 'stack',
         title = 'Data Coverage of Watercraft Complexity Data Fields'
  )
```