---
title: "2025 Veliger Sampling Locations"
author: "John Phelan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = F, warning = F, message = F)
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(tidyverse)
library(sf)
library(bcdata)
library(bcmaps)
library(readxl)
library(leaflet)
library(leafpop)
library(leaflet.extras)
library(janitor)
library(here)
library(openxlsx)
library(htmlwidgets)

lan_root = "//SFP.IDIR.BCGOV/S140/S40203/WFC AEB/General/2 SCIENCE - Invasives/AIS_R_Projects/CMadsen_Wdrive/shared_data_sets/"


dat<-openxlsx::read.xlsx(here::here("./data/Copy of BC Veliger Sampling Inventory 2025.xlsx"))


```


```{r data grab}

dat <- dat |> 
  clean_names()

dat<- dat |> 
  rename(count =  x1)

dat<- dat |> 
  mutate(date_collected_yyyy_mm_dd = as.Date(as.numeric(date_collected_yyyy_mm_dd), origin = "1899-01-01"))

dat_complete<-dat |> 
  filter(!is.na(lat_decimal_degrees) | !is.na(long_decimal_degrees))
dat_missing<-dat |> 
  filter(is.na(lat_decimal_degrees) | is.na(long_decimal_degrees))


#fix wrong longs - all should be negative
dat_complete<-dat_complete |> 
  mutate(long_decimal_degrees = ifelse(long_decimal_degrees > 0, long_decimal_degrees * -1, long_decimal_degrees))


dat_sf<-st_as_sf(dat_complete, coords = c("long_decimal_degrees", "lat_decimal_degrees"), crs = 4326)

```

```{r First Nations Polygons}

firstNationsPIP<-st_read(paste0(lan_root, "FN_PIP_Consultation_Areas.shp"), quiet = T)

firstNationsPIP<- st_transform(firstNationsPIP, crs = 4326)
firstNationsPIP <- st_make_valid(firstNationsPIP)
# ggplot()+
#   geom_sf(data = firstNationsPIP)

```


```{r manipulations}

#lets only keep Nations that overlap with where sampling has occurred?
Nations_involved <- st_filter(firstNationsPIP, dat_sf, .predicate = st_intersects)
  
Nations_involved <- Nations_involved |> 
  rename(Nation = CNSLTN_A_1)
# ggplot() +
#   geom_sf(data = Nations_involved, fill = "lightblue", color = "black") +
#   geom_sf(data = dat_sf, color = "red")

overlap_list <- st_intersects(Nations_involved)

# Create a column that lists all overlapping nations for each polygon
Nations_involved$all_overlapping_nations <- sapply(seq_along(overlap_list), function(i) {
  inds <- overlap_list[[i]]
  paste(unique(Nations_involved$Nation[inds]), collapse = ", ")
})


###############
# --------------- These polygons below do not help - too small and poor coverage
##############


# ls<-bcdc_list()
# 
# ls_filtered <- ls[str_detect(ls, regex("nations", ignore_case = TRUE))]
# 
# ls_filtered
# 
# nations_lands<- bcdc_get_data("first-nations-treaty-lands")
# 
# nations_lands<- st_transform(nations_lands, 4326)
# # small amount of data
# #plot(st_geometry(nations_lands))
# 
# nations_areas<- bcdc_get_data("first-nations-treaty-areas")
# 
# nations_areas<- st_transform(nations_areas, 4326)
# small amount of data
# plot(st_geometry(nations_areas))
# no overlap
# Nations_involved <- st_filter(nations_areas, dat_sf, .predicate = st_intersects)



```

## Overlap of Veliger Sampling and First Nations territories

From the veliger sampling data, `r nrow(dat_complete)` samples were returned with latitude and longitude, and `r nrow(dat_missing)` were without co-ordinates.

From the [Profiles of Indigenous Peoples (PIP)](https://catalogue.data.gov.bc.ca/dataset/profiles-of-indigenous-peoples-pip-consultation-areas-public-map-service), there are a total of `r nrow(Nations_involved)` First Nations with territories that overlap with the 2025 Zebra/Quagga veliger sampling being conucted in BC.

```{r leaflet map, fig.width=10, fig.height=8}

leaflet() |> 
  addTiles(group = "OSM") |> 
  addProviderTiles(providers$CartoDB.Positron, group = "CartoDB") |> 
  addLayersControl(
    position = 'bottomleft',
    baseGroups = c("CartoDB","OSM"),
    options = layersControlOptions(collapsed = FALSE)
  ) |> 
  addPolygons(
    data = Nations_involved,
    color = "black",
    fillColor = "lightblue",
    fillOpacity = 0.5,  # transparent to see overlaps
    weight = 1,
    label = ~Nation,  # hover label
    popup = leafpop::popupTable(
      Nations_involved,
      zcol = "all_overlapping_nations"
    )
  ) |> 
  addCircleMarkers(
    data = dat_sf,
    color = "purple",
    fillOpacity = 0.8,
    radius = 2,
    label = ~paste0(waterbody, ", ", date_collected_yyyy_mm_dd),
    popup = leafpop::popupTable(
      dat_sf,
      zcol = c(
        "waterbody",
        "location_site_name",
        "date_collected_yyyy_mm_dd",
        "secchi_depth_m",
        "water_temperature_c",
        "p_h_in_water_column",
        "zebra_quagga_mussels_veligers",
        "sampling_group_agency"
      )
    )
  )


```



