---
title: "IMDP Crossyear Comparison"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: lumen
    highlight: github
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
library(tidyverse)
rm(list = ls())
```

## IMDP Cross-year Comparison

This R Markdown document compares data from watercraft inspections carried out as part of the Invasive Mussel Defence Program (IMDP). 

The following comparisons between years are generated:

1. High Risk Inspections by Station 
2. Decontamination Orders Issued by Station
3. Decontaminations Performed by Station
4. Previously Inspected by Station
5. Commercially Hauled by Station
6. Previous Inspection/Referral of Scheduled Inspections
7. Percent Previously Inspected for Boat Launch Records
8. Number of Previously Inspected (or not) High-risk Inspections
9. Hwy 97C - Boat Source Jurisdiction
10. Hwy 97C - Number and Source of Previously Inspected Craft

Please note: for best results, download the 2021 .csv file summary of all of the data from metabase and place it in the data folder inside the LocalRWork folder.


```{r load_in_data, include=FALSE, cache=TRUE}
#Look for all .csv files in the folder "IMDP_metabase" - same their names.
myfiles = list.files("data/IMDP_metabase/", 
           pattern = ".csv", full.names = T)

#Read in each .csv file that was found above. All columns read as character data.
dat = lapply(myfiles, read.csv, colClasses = "character")

#Combine the .csv files into one big table.
dat = dat %>% 
  bind_rows()

#Read in 2019 data for a couple figures.
dat2019 = read.csv("C:/Users/CMADSEN/Downloads/LocalRWork/output/IMDP_2019_HighRisk_DeconOrdIss_DeconPerf.csv",
                   colClasses = "character")
  

#Do some obligatory data cleaning to remove test records, correct some things.
dat = dat %>% 
  filter(duplicated(raw_timestamp) != T,
         str_detect(Shift.Start.Comment, "TEST") == F,
         str_detect(Shift.Start.Comment, "Test") == F) %>% 
  mutate(Station = replace(Station, str_detect(Station, "Scheduled"), "Scheduled Inspection")) %>% 
  mutate(Station = replace(Station, str_detect(Station, "Olsen"), "Olsen (Hwy 3)")) %>% 
  mutate(Station = replace(Station, str_detect(Station, "Pacific"), "Pacific")) %>% 
  mutate(Year = lubridate::year(as.Date(raw_timestamp))) %>% 
  mutate(Year = factor(Year,levels = c(2019,2020,2021)))

dat2019 = dat2019 %>% 
  mutate(Station = replace(Station, str_detect(Station, "Schedule"), "Scheduled Inspection")) %>% 
  mutate(Station = replace(Station, str_detect(Station, "Olsen"), "Olsen (Hwy 3)")) %>% 
  mutate(Station = replace(Station, Station == "Pacific Surrey", "Pacific"),
         Station = replace(Station, Station == "Pacific Border", "Pacific")) %>% 
  filter(!Station %in% c("Scheduled Inspection","Cascade Border","Kamloops Road Check","Sumas Border")) %>% 
  rename(Watercraft.Risk.Assessment.ID = Watercraft_Risk_Assessment_ID,
         High.Risk.AIS.Ind = High_Risk_AIS_Ind,
         Shift.Start.Comment = Shift_Start_Comment,
         Decontamination.order.issued.Ind = Decontamination_order_issued_Ind,
         Decontamination.Performed.Ind = Decontamination_Performed_Ind)

#Save the complete dataset and put it aside - we'll use it below for just a couple of figures.
dat_all = dat

#Filter out stations that are not central.
dat = dat %>% 
  filter(!Station %in% c("Penticton Roving","Hwy 97C","Keremeos (Hwy 3)","Greenwood","Kaleden","Other","Scheduled Inspection","Midway","Boat Launch - Okanagan"))
```

Comparison 
Martina asked for the maximum number of inspections in a day for each station in 2021.

```{r max_insp_2021, error=TRUE}
MaxInspDaily_Station = dat %>% 
  filter(Year == "2021") %>% 
  mutate(MyDay = str_extract(Inspection.Date,"^[0-9]{2}"),
         MyMonth = str_remove_all(str_remove_all(Inspection.Date, "[0-9]*"),"-")) %>%
  group_by(Station,MyMonth,MyDay) %>% 
  summarise(DailyCount = n()) %>%
  group_by(Station) %>% 
  mutate(MaxCount = max(DailyCount)) %>% 
  dplyr::select(Station, MaxCount) %>% 
  distinct() %>% 
  arrange(desc(MaxCount))

openxlsx::write.xlsx(MaxInspDaily_Station,"I:/SPECIES/Zebra_Quagga_Mussel/Operations/Watercraft Inspection Data/2021 data/Max Inspections in a day by station for 2021 data.xlsx")
```


Martina also asked to see all of the records that were scheduled inspections. This excel file is saved to SPECIES -> Zebra_Quagga_Mussel -> Operations -> Watercraft Inspection Data -> 2021 data folder.

```{r scheduled_insp, echo = T, error=TRUE}
SchedInsp = dat_all %>%
  filter(Year == "2021") %>% 
  filter(str_detect(Station, "Scheduled"))

openxlsx::write.xlsx(SchedInsp, 
"I:/SPECIES/Zebra_Quagga_Mussel/Operations/Watercraft Inspection Data/2021 data/Scheduled Inspections 2021.xlsx",
overwrite = TRUE)
```

# High Risk Inspections by Station


```{r high_risk, echo=FALSE}
highriskdat = dat %>% 
  bind_rows(dat2019 %>% 
              mutate(Year = 2019) %>% 
              mutate(Year = factor(Year, levels = c(2019,2020,2021)))) %>% 
  filter(High.Risk.AIS.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(Count = n()) %>% 
  mutate(TotalStationCount = sum(Count)) %>% 
  ungroup() %>% 
  arrange(desc(TotalStationCount))

highriskdat$Station = factor(highriskdat$Station, 
                             levels = unique(highriskdat$Station))  

highriskdat %>% 
  ggplot() + 
  geom_col(aes(x = Station, y = Count, fill = Year),
           position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Station, y = Count+15), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of High-risk Inspections") +
  #scale_y_continuous(breaks = c(0,15,30,60,90,120)) +
  scale_x_discrete() +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


# Decontamination Orders Issued by Station


```{r decon_orders, echo=FALSE}
decondat = dat %>% 
  bind_rows(dat2019 %>% 
              mutate(Year = 2019) %>% 
              mutate(Year = factor(Year, levels = c(2019,2020,2021)))) %>% 
  filter(Decontamination.order.issued.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(Count = n()) %>% 
  mutate(TotalStationCount = sum(Count)) %>% 
  ungroup() %>% 
  arrange(desc(TotalStationCount))

decondat$Station = factor(decondat$Station, 
                             levels = unique(decondat$Station))  

decondat %>% 
  ggplot() + 
  geom_col(aes(x = Station, y = Count, fill = Year), 
           position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Station, y = Count+3), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of Decontamination Orders Issued") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```

# Decontaminations Performed by Station	


```{r decon_performed, echo=FALSE}
decondat = dat %>%
  bind_rows(dat2019 %>% 
              mutate(Year = 2019) %>% 
              mutate(Year = factor(Year, levels = c(2019,2020,2021)))) %>% 
  filter(Decontamination.Performed.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(Count = n()) %>% 
  mutate(TotalStationCount = sum(Count)) %>% 
  ungroup() %>% 
  arrange(desc(TotalStationCount))

decondat$Station = factor(decondat$Station, 
                             levels = unique(decondat$Station))  

decondat %>% 
  ggplot() + 
  geom_col(aes(x = Station, y = Count, fill = Year), 
           position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Station, y = Count+3), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of Decontaminations Performed") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


# Extra Figure Just for 2021
## Number of Decontamination Orders Issued Vs. Performed

```{r decon_comp,echo=FALSE}
decondat = dat %>%
  filter(Year == 2021) %>% 
  filter(is.na(Decontamination.Reference) != T,
         Decontamination.Reference != "None",
         Decontamination.Reference != "-1",
         Decontamination.Reference != '') %>% 
  group_by(Station,Year) %>% 
  summarise(Referenced = n()) %>% 
  left_join(dat %>%
  filter(Year == 2021) %>% 
  filter(Decontamination.Performed.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(Performed = n())) %>% 
  pivot_longer(cols = c(Referenced,Performed)) %>% 
  arrange(desc(value))

decondat$Station = factor(decondat$Station, 
                             levels = unique(decondat$Station))  

decondat %>% 
  ggplot() + 
  geom_col(aes(x = Station, y = value, fill = name), 
           position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = value, x = Station, y = value+2), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Decontamination Referenced/Performed", fill = "Decontamination") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```

#	Previously Inspected by Station


```{r prev_insp, echo=FALSE}
prevdat = dat %>%
  filter(Previous.Inspection.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(PreviousInspectedCount = n()) %>% 
  left_join(
  dat %>% 
  group_by(Station,Year) %>% 
  summarise(TotalInspectedCount = n())
  ) %>% 
  mutate(Perc_Prev_Insp = round(100*(PreviousInspectedCount/(TotalInspectedCount)),2)) %>% 
  arrange(desc(Perc_Prev_Insp))

prevdat$Station = factor(prevdat$Station, 
                             levels = unique(prevdat$Station))  

prevdat %>% 
  ggplot(aes(x = Station, y = Perc_Prev_Insp, fill = Year)) + 
  geom_col(position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = round(Perc_Prev_Insp,1), x = Station, y = Perc_Prev_Insp+2), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Inspections w/ Previous Inspections (%)") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


#	Commercially Hauled by Station


```{r comm_haul, echo=FALSE}
commdat = dat %>% 
  filter(Commercially.Hauled.Ind == "true") %>% 
  group_by(Station,Year) %>% 
  summarise(Count = n()) %>% 
  mutate(TotalStationCount = sum(Count)) %>% 
  ungroup() %>% 
  arrange(desc(Count))

commdat$Station = factor(commdat$Station, 
                             levels = unique(commdat$Station))  

commdat %>% 
  ggplot(aes(x = Station, y = Count, fill = Year)) + 
  geom_col(position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Station, y = Count+5), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of Commercially Hauled Inspections") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


# Previous Inspection/Referral of Scheduled Inspections


```{r source_scheduled_inspec, echo=FALSE,fig.width = 14,fig.height=8}
schedinsp_source = dat_all %>% 
  filter(str_detect(Station, "cheduled Inspection.*")) %>% 
  rename(Source = Previous.Inspection.Source.Name) %>% 
  mutate(Source = replace(Source, Source == "", "Data Missing")) %>% 
  filter(Source != "Other - BC") %>% 
  group_by(Year,Source) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count)) %>% 
  ungroup()
  
schedinsp_source$Source = factor(schedinsp_source$Source, 
                             levels = unique(schedinsp_source$Source))  

schedinsp_source %>% 
  ggplot(aes(x = Source, y = Count, fill = Year)) + 
  geom_col(position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Source, y = Count+2), 
            colour = "black",
            size = 5,
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Previous Inspection/Referral") +
  theme_minimal() +
  scale_y_continuous(limits = c(0, 40)) +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 18),
        axis.text.x = element_text(size=18,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


# Percent Previously Inspected for Boat Launch Records


```{r perc_prev_insp, echo = FALSE}
previnsp = dat_all %>%
  filter(Station == "Boat Launch - Okanagan") %>% 
  filter(Previous.Inspection.Ind == "true") %>% 
  group_by(Year,Previous.Inspection.Source.Name) %>% 
  summarise(PreviousInspectedCount = n()) %>% 
  left_join(
  dat_all %>% 
  filter(Station == "Boat Launch - Okanagan") %>% 
  group_by(Year) %>% 
  summarise(TotalInspectedCount = n())
  ) %>% 
  mutate(Perc_Prev_Insp = round(100*(PreviousInspectedCount/(TotalInspectedCount)),2)) %>% 
  arrange(desc(Perc_Prev_Insp))

previnsp$Previous.Inspection.Source.Name = factor(previnsp$Previous.Inspection.Source.Name, 
                             levels = unique(previnsp$Previous.Inspection.Source.Name))  

previnsp %>% 
  ggplot(aes(x = Previous.Inspection.Source.Name, y = Perc_Prev_Insp, fill = Year)) + 
  geom_col(position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = round(Perc_Prev_Insp,1), x = Previous.Inspection.Source.Name, y = Perc_Prev_Insp+1), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Proportion of Previous Inspections (%)",
       x = "Previous Inspection Location",
       title = "Boat Launch - Okanagan") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```



# Number of Previously Inspected (or not) High-risk Inspections


```{r insp_or_not_high_risk, echo=FALSE}
highrisk_insp_or_not = dat_all %>% 
  filter(Station == "Boat Launch - Okanagan") %>% 
  filter(High.Risk.AIS.Ind == "true") %>% 
  group_by(Year,Previous.Inspection.Ind) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count))

highrisk_insp_or_not %>% 
  ggplot(aes(x = Previous.Inspection.Ind, y = Count, fill = Previous.Inspection.Ind)) + 
  geom_col(position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Previous.Inspection.Ind, y = Count+1), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of Previously Inspected Watercraft", x = "Previous Inspection", title = "Boat Launch - Okanagan") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```

# Hwy 97C - Boat Source Jurisdiction


```{r hwy97_boat_source, echo=FALSE}
hwy97_boatsource = dat_all %>% 
  filter(Station == "Hwy 97C") %>% 
  group_by(Station,Province.Code) %>% 
  summarise(Count = n()) %>% 
  arrange(desc(Count))

hwy97_boatsource$Province.Code = factor(hwy97_boatsource$Province.Code, 
                             levels = unique(hwy97_boatsource$Province.Code))  

hwy97_boatsource %>% 
  ggplot() + 
  geom_col(aes(x = Province.Code, y = Count, fill = Station), 
           position = position_dodge2(preserve = "single")) +
  geom_text(aes(label = Count, x = Province.Code, y = Count+20), 
            colour = "black",
            position = position_dodge2(1, preserve = "single")) +
  labs(y = "Number of Inspections", x = "Boat Source Jurisdiction") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        text = element_text(size = 13),
        axis.text.x = element_text(size=13,angle = 60,hjust=1)) +
  scale_fill_brewer(palette = "Set1")
```


```{r hwy97_prev_craft, include = FALSE, echo=FALSE,fig.width = 14,fig.height=8}
dat_all %>% 
  filter(Station == "Hwy 97C") %>% 
  group_by(Previous.Inspection.Ind) %>% 
  summarise(Numba = n()) 

dat_all %>% 
  filter(Station == "Hwy 97C") %>% 
  filter(High.Risk.AIS.Ind == "true") %>% 
  select(Previous.Waterbody.1.Name)
# hwy97_prevcraft = dat_all %>% 
#   rename(insp_name = Previous.Inspection.Source.Name) %>% 
#   filter(Station == "Hwy 97C",
#          Previous.Inspection.Ind == "true") %>% 
#   group_by(Station,insp_name) %>% 
#   summarise(Count = n()) %>% 
#   arrange(desc(Count)) %>% 
#   slice(1:15)

# hwy97_prevcraft$insp_name = factor(hwy97_prevcraft$insp_name, 
#                              levels = unique(hwy97_prevcraft$insp_name))  
# 
# hwy97_prevcraft %>% 
#   ggplot() + 
#   geom_col(aes(x = insp_name, y = Count, fill = Station),
#            position = position_dodge2(preserve = "single")) +
#   geom_text(aes(label = Count, x = insp_name, y = Count+3), 
#             colour = "black",
#             size = 5,
#             position = position_dodge2(1, preserve = "single")) +
#   labs(y = "Number of Inspections", x = "Location of Previous Inspection") +
#   theme_minimal() +
#   theme(panel.grid.minor = element_blank(),
#         text = element_text(size = 18),
#         axis.text.x = element_text(size=16,angle = 60,hjust=1)) #+
#  scale_fill_brewer(palette = "Set1")
```