---
title: "Search for Mussel Fouled Boats"
author: "John Phelan"
date: "2026-02-26"
output: html_document
---

```{r set_up, echo = F, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)



  source(here::here("02_IMDP_Figure_Generator/utils/initialise_plots.R"))

#What is the sheet name of the destination regions for the mussel fouled tracking sheet?
if(my.year != 2024){
  MF_tracker_sheet = "DestRegions"
}
if(my.year == 2024){
  MF_tracker_sheet = "Final List"
}


source(here::here("02_IMDP_Figure_Generator", "utils", "load_figure_data.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "station_manipulation.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "dat_to_stations.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "compliance_dat.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "compliance_total_data.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "blowbys_by_hour.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "high_risk_by_station.R"))



```



```{r fix_data}


```


## Data Exploration {.tabset}

### Suspect ID's

ID's being inspected are: 

- 150651 - This record had no notes in the database that would suggest it was mussel fouled.


### Comment Searching

For all stations we search for any comments mentioning: mussel - fouled - zebra - ZQ - dreissinid

```{r comment_search}

```






### Initial Data Exploration - Mount Robson

Initial counts and data exploration - short summary of what we have for Mount Robson.

There were no blow bys recorded in the main table - I might need to look elsewhere for these.

```{r get_data}
if (!exists('dat')) {
  dat <- read.xlsx(paste0(my.data.folder, '/figure_dat.xlsx')) |> 
    mutate(TimeOfInspection = convertToDateTime(TimeOfInspection)) |> 
    as_tibble() #|> 
    # mutate(
    #   Station = case_when(
    #     Station == "Penticton Roving" ~ "Penticton",
    #     Station == "Lower Mainland Roving" ~ "Lower Mainland",
    #     TRUE ~ Station
    #   )
    # )
}


mt_robson = dat |> 
  filter(str_detect(Station, "Robson"))
# what are the comments 
unique(mt_robson$Shift_Start_Comment)

tbl_mussels <- mt_robson |> 
  count(MusselsFound_Ind, name = "Mussels Found")

tbl_mussels |>
  kable(caption = "Mussels Found - Mount Robson") |>
  kable_styling(full_width = FALSE)

tbl_highrisk <- mt_robson |> 
  count(High_Risk_AIS_Ind, name = "High Risk")

tbl_highrisk |>
  kable(caption = "High Risk Watercraft - Mount Robson") |>
  kable_styling(full_width = FALSE)

tbl_drainplug <- mt_robson |> 
  count(Watercraft_Has_Drainplugs_Ind, name = "Watercraft has Drainplug")

tbl_drainplug |>
  kable(caption = "Watercraft with Drainplug") |>
  kable_styling(full_width = FALSE)

tbl_removed <- mt_robson |> 
  filter(Watercraft_Has_Drainplugs_Ind == TRUE) |> 
  count(Drainplug_Removed_at_Inspection_Ind, name = "Drainplugs Pulled")

tbl_removed |>
  kable(caption = "Drainplugs Removed at Inspection") |>
  kable_styling(full_width = FALSE)

tbl_previous <- mt_robson |>
  count(Previous_Waterbody_1_Province_Or_State,
        name = "Previous Province or State") |>
  arrange(desc(`Previous Province or State`))

tbl_previous |>
  kable(caption = "Previous Province or State") |>
  kable_styling(full_width = FALSE)

tbl_destination <- mt_robson |>
  count(Destination_Waterbody_1_Province_Or_State,
        name = "Destination Province or State") |>
  arrange(desc(`Destination Province or State`))

tbl_destination |>
  kable(caption = "Destination Province or State") |>
  kable_styling(full_width = FALSE)

tbl_complexity <- mt_robson |>
  select(
    Non_Motorized_Counter,
    Simple_Counter,
    Complex_Counter,
    Very_Complex_Counter
  ) |>
  pivot_longer(everything(),
               names_to = "variable",
               values_to = "value") |>
  group_by(variable, value) |>
  summarise(n = n(), .groups = "drop")

tbl_complexity |>
  kable(caption = "Watercraft Complexity Summary") |>
  kable_styling(full_width = FALSE)

# tbl_blowby <- mt_robson |> 
#   select(Non_Motorized_Blow_Bys_Counter,
#          Motorized_Blow_Bys_Counter,
#          Total_BlowBys)|>
#   dplyr::mutate(
#     dplyr::across(dplyr::everything(), as.numeric)
#   ) |> 
#   pivot_longer(everything(),
#                names_to = "variable",
#                values_to = "value") |>
#   group_by(variable, value) |>
#   summarise(n = n(), .groups = "drop")
# 
# tbl_blowby |> 
#   kable(caption = "Blowbys") |>
#   kable_styling(full_width = FALSE)

tbl_decon <- mt_robson |> 
  count(Decontamination_Performed_Ind, name = "Decon performed")

tbl_decon |>
  kable(caption = "Decon performed - Mount Robson") |>
  kable_styling(full_width = FALSE)

tbl_quar <- mt_robson |> 
  count(Quarantine_Period_Issued_Ind, name = "Quarantine order issued")

tbl_quar |>
  kable(caption = "Quarantine issued - Mount Robson") |>
  kable_styling(full_width = FALSE)

tbl_dress <- mt_robson |> 
  count(Dreissenid_mussels_found_during_previous_inspection, name = "Dreissenid found previously")

tbl_dress |>
  kable(caption = "Dreissenid found previously - Mount Robson") |>
  kable_styling(full_width = FALSE)



```


### 
