---
title: "Search for Mussel Fouled Boats"
author: "John Phelan"
date: "2026-02-26"
output: html_document
---

```{r set_up, echo = F, include=FALSE, message = F, warning = F}
knitr::opts_chunk$set(echo = F, message = F, warning = F)


library(kableExtra)
library(DT)

source(here::here("02_IMDP_Figure_Generator/utils/initialise_plots.R"))

#What is the sheet name of the destination regions for the mussel fouled tracking sheet?
if(my.year != 2024){
  MF_tracker_sheet = "DestRegions"
}
if(my.year == 2024){
  MF_tracker_sheet = "Final List"
}


source(here::here("02_IMDP_Figure_Generator", "utils", "load_figure_data.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "station_manipulation.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "dat_to_stations.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "compliance_dat.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "compliance_total_data.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "blowbys_by_hour.R"))

source(here::here("02_IMDP_Figure_Generator", "utils", "high_risk_by_station.R"))



```



```{r fix_data}

if (!exists('dat')) {
  dat <- read.xlsx(paste0(my.data.folder, '/figure_dat.xlsx')) |> 
    mutate(TimeOfInspection = convertToDateTime(TimeOfInspection)) |> 
    as_tibble() #|> 
  # mutate(
  #   Station = case_when(
  #     Station == "Penticton Roving" ~ "Penticton",
  #     Station == "Lower Mainland Roving" ~ "Lower Mainland",
  #     TRUE ~ Station
  #   )
  # )
}


```


## Data Exploration {.tabset}

### Watercraft Inspection ID to be Investigated

150651 was explored for any inconsistencies and there was nothing in the database to suggest that this was mussel fouled.

### Comment Exploration


To explore the comments, we look for: mussel - fouled - zebra - ZQ - dreissinid. Any instances of these will return the comments. There are no mentions of these in the start or end comments. 6 entries were found for 2025 with these words in the comments. 

135441 had high risk indicated. It had marine mussels (barnacles) found on inspection. It originated from a high risk area. It was previously in dry storage. Decon order issued (2042). Seal was issued (6875).  "No decontamination - pressure washer not working." 


```{r explore comments}


keywords <- c("mussel", "fouled", "zebra", "ZQ", "dreissinid", "quagga")

# dat |>
#   filter(str_detect(
#     Shift_Start_Comment,
#     regex(paste(keywords, collapse = "|"), ignore_case = TRUE)
#   ))
# 
# 
# dat |>
#   filter(str_detect(
#     Shift_End_Comment,
#     regex(paste(keywords, collapse = "|"), ignore_case = TRUE)
#   ))


dat |>
  mutate(General_Comment = as.character(General_Comment)) |>
  filter(str_detect(
    General_Comment,
    regex(paste(keywords, collapse = "|"), ignore_case = TRUE)
  )) |>
  # select(Watercraft_Risk_Assessment_ID, MusselsFound_Ind, High_Risk_AIS_Ind, 
  #        Station, General_Comment, Shift_Start_Comment, Shift_End_Comment, Watercraft_Has_Drainplugs_Ind, Drainplug_Removed_at_Inspection_Ind, Decontamination_Performed_Ind, Quarantine_Period_Issued_Ind, Dreissenid_mussels_found_during_previous_inspection, Clean_Drain_Dry_After_Inspection_Ind, Seal_Issued_Ind, DryStorage, ) |> 
  datatable(
    options = list(
      pageLength = 10,
      autoWidth = TRUE
    ),
    rownames = FALSE
  )




```






```{r get_data}

# ### Initial Data Exploration - Mount Robson
# 
# Initial counts and data exploration - short summary of what we have for Mount Robson.
# 
# There were no blow bys recorded in the main table - I might need to look elsewhere for these.
# 
# mt_robson = dat |> 
#   filter(str_detect(Station, "Robson"))
# # what are the comments 
# unique(mt_robson$Shift_Start_Comment)
# 
# tbl_mussels <- mt_robson |> 
#   count(MusselsFound_Ind, name = "Mussels Found")
# 
# tbl_mussels |>
#   kable(caption = "Mussels Found - Mount Robson") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_highrisk <- mt_robson |> 
#   count(High_Risk_AIS_Ind, name = "High Risk")
# 
# tbl_highrisk |>
#   kable(caption = "High Risk Watercraft - Mount Robson") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_drainplug <- mt_robson |> 
#   count(Watercraft_Has_Drainplugs_Ind, name = "Watercraft has Drainplug")
# 
# tbl_drainplug |>
#   kable(caption = "Watercraft with Drainplug") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_removed <- mt_robson |> 
#   filter(Watercraft_Has_Drainplugs_Ind == TRUE) |> 
#   count(Drainplug_Removed_at_Inspection_Ind, name = "Drainplugs Pulled")
# 
# tbl_removed |>
#   kable(caption = "Drainplugs Removed at Inspection") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_previous <- mt_robson |>
#   count(Previous_Waterbody_1_Province_Or_State,
#         name = "Previous Province or State") |>
#   arrange(desc(`Previous Province or State`))
# 
# tbl_previous |>
#   kable(caption = "Previous Province or State") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_destination <- mt_robson |>
#   count(Destination_Waterbody_1_Province_Or_State,
#         name = "Destination Province or State") |>
#   arrange(desc(`Destination Province or State`))
# 
# tbl_destination |>
#   kable(caption = "Destination Province or State") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_complexity <- mt_robson |>
#   select(
#     Non_Motorized_Counter,
#     Simple_Counter,
#     Complex_Counter,
#     Very_Complex_Counter
#   ) |>
#   pivot_longer(everything(),
#                names_to = "variable",
#                values_to = "value") |>
#   group_by(variable, value) |>
#   summarise(n = n(), .groups = "drop")
# 
# tbl_complexity |>
#   kable(caption = "Watercraft Complexity Summary") |>
#   kable_styling(full_width = FALSE)
# 
# # tbl_blowby <- mt_robson |> 
# #   select(Non_Motorized_Blow_Bys_Counter,
# #          Motorized_Blow_Bys_Counter,
# #          Total_BlowBys)|>
# #   dplyr::mutate(
# #     dplyr::across(dplyr::everything(), as.numeric)
# #   ) |> 
# #   pivot_longer(everything(),
# #                names_to = "variable",
# #                values_to = "value") |>
# #   group_by(variable, value) |>
# #   summarise(n = n(), .groups = "drop")
# # 
# # tbl_blowby |> 
# #   kable(caption = "Blowbys") |>
# #   kable_styling(full_width = FALSE)
# 
# tbl_decon <- mt_robson |> 
#   count(Decontamination_Performed_Ind, name = "Decon performed")
# 
# tbl_decon |>
#   kable(caption = "Decon performed - Mount Robson") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_quar <- mt_robson |> 
#   count(Quarantine_Period_Issued_Ind, name = "Quarantine order issued")
# 
# tbl_quar |>
#   kable(caption = "Quarantine issued - Mount Robson") |>
#   kable_styling(full_width = FALSE)
# 
# tbl_dress <- mt_robson |> 
#   count(Dreissenid_mussels_found_during_previous_inspection, name = "Dreissenid found previously")
# 
# tbl_dress |>
#   kable(caption = "Dreissenid found previously - Mount Robson") |>
#   kable_styling(full_width = FALSE)



```

