---
title: "ZQ Mussel and Golden Mussel Calcium Tolerance Comparison"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
---

# Sections {.tabset}

## Cleaning 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
# Load libraries
library(terra)
library(tidyverse)
library(lubridate)
library(sf)
library(ggspatial)
library(ggthemes)
library(bcdata)
library(DBI)
library(RSQLite)
library(tidyhydat)
library(patchwork)
```

```{r read_and_set_options}
my_opts = read_csv(paste0(here::here(),"/Options.csv")) %>% 
  as.data.frame()

my.year = my_opts$year

temp_limit_zqm = my_opts$zqm_risk_temp_limit
temp_limit_golden = 15
climate_model = my_opts$zqm_risk_climate_model

# setwd(paste0(my_opts$base_dir,"01_DataCleaning/"))
mywd = getwd()
# Make path to 01_DataCleaning/data

data.path = paste0(stringr::str_remove(mywd, "04_.*"),"/01_DataCleaning/data/")

#Create organizational folders in your working directory, if they don't exist.
#data...
# if(!dir.exists(paste0(getwd(),"/data/"))) {
#   dir.create(paste0(getwd(),"/data/"))
# }

#output...
if(!dir.exists(paste0(getwd(),"/output/"))) {
  dir.create(paste0(getwd(),"/output/"))
}
```

```{r load_in_shapefiles}
# BC shapefile
bc_shp = bcmaps::bc_bound() %>% 
  st_transform(crs = 4326) |> 
  dplyr::summarise()

# WorldClim future temperature raster data. 
if(my_opts$zqm_risk_climate_model == "ssp245"){
  tempr <- rast(paste0(data.path,"wc2.1_2.5m_bioc_MRI-ESM2-0_ssp245_2021-2040.tif"), lyrs = 10)
  tempr = mask(crop(tempr, bc_shp), vect(bc_shp))
} else if(my_opts$zqm_risk_climate_model == "ssp370"){
  tempr <- rast(paste0(data.path,"wc2.1_2.5m_bioc_MRI-ESM2-0_ssp370_2021-2040.tif"), lyrs = 10)
  tempr = mask(crop(tempr, bc_shp), vect(bc_shp))
}

# Query from EMS!
ems_path = paste0(stringr::str_extract(my_opts$base_dir,".*long_term_projects"),"/EMS/output/EMS.sqlite")
conn = dbConnect(SQLite(), ems_path)
df = DBI::dbGetQuery(conn, "SELECT * FROM results where PARAMETER like 'Calcium %' or PARAMETER like 'pH'")
# ems_temp = DBI::dbGetQuery(conn, "SELECT * FROM results where PARAMETER like 'Temperature%'")
DBI::dbDisconnect(conn)

#Get rid of columns 29 to 60, which we do not need.
df = df[,c(1:28)]

```

```{r combine_data}
df = df |> 
  dplyr::select(EMS_ID, MONITORING_LOCATION,LOCATION_TYPE,COLLECTION_START,COLLECTION_END,SAMPLING_AGENCY,COLLECTION_METHOD,PARAMETER,ANALYTICAL_METHOD,RESULT,UNIT,LONGITUDE,LATITUDE)

# Ensure we don't have points with missing locations or sampling dates.
df = df |> 
  dplyr::filter(!is.na(LATITUDE),
                !is.na(LONGITUDE)) |> 
  dplyr::filter(!(is.na(COLLECTION_START) & is.na(COLLECTION_END)))

# Clean up date columns!
df = df |> 
  dplyr::mutate(COLLECTION_START = lubridate::ymd_hms(COLLECTION_START),
                COLLECTION_END = lubridate::ymd_hms(COLLECTION_END))

hist(lubridate::year(df$COLLECTION_END)) #What's the distribution of data through time?
```

```{r filter_outliers_from_calcium_and_pH}
df_f = df |> 
  dplyr::filter(!is.na(RESULT)) |> 
  mutate(retain_point = case_when(
    PARAMETER == 'pH' & (RESULT <= 0 | RESULT > 16.5) ~ F,
    PARAMETER != 'pH' & (RESULT <= 0 | RESULT > 300) ~ F,
    T ~ T
  )) |> 
  dplyr::filter(retain_point)
```

```{r read_in_wb_and_subw_shapefiles}
#Read in the water body shapefiles
waterbodies = read_sf(paste0(data.path,"summarized_bc_waterbodies_same_gnis_joined.gpkg"))
# Read in subwatershed (N = 246) shapefile
watershed = read_sf(paste0(my_opts$remote_spatial_data,"shared_data_sets/WatershedGroups_lowres.shp"))
```

```{r match_EMS_records_to_wb}
# ---------------
# We have a decision to make here: do we filter EMS data by whether or not it matches to a waterbody (either by name or spatially)?

# Calculate number of data points for pH and calcium
total_pH = nrow(df_f |> dplyr::filter(PARAMETER == 'pH'))

total_calc = nrow(df_f |> dplyr::filter(PARAMETER != 'pH'))

# Simple spatial match to waterbodies

df_f_match_to_wbs = df_f |>
  dplyr::filter(!is.na(LATITUDE)) |> 
  st_as_sf(coords = c('LONGITUDE','LATITUDE'),
           crs = 4326) |> 
  sf::st_transform(crs = 3005) |> 
  st_join(waterbodies)

df_f_match_to_wbs = df_f_match_to_wbs |>
  dplyr::filter(!is.na(WATERSH))

pH = df_f_match_to_wbs |> filter(PARAMETER == 'pH') |> dplyr::rename(WATERSHED_ = WATERSH)

calc = df_f_match_to_wbs |> filter(PARAMETER != 'pH') |> dplyr::rename(WATERSHED_ = WATERSH)

# # Summarized number of data points for pH and calcium by subwatershed.
pH_by_subw = pH |>
  sf::st_drop_geometry() |>
  dplyr::count(WATERSHED_)

calc_by_subw = calc |>
  sf::st_drop_geometry() |>
  dplyr::count(WATERSHED_)

```

## Number of Data Points

```{r visualize_number_pH_data_points_per_subw}
# Add summaries to subwatershed; plot.
watershed |> 
  dplyr::left_join(
    pH_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = log(n))) + 
  labs(title = 'pH Readings from EMS across BC',
       subtitle = paste0(nrow(pH_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(pH)/total_pH),1),'% of ~',round(total_pH/1000,0),'K points matched a waterbody'),
       fill = 'Natural log\nof number of \ndata points',
       caption = 'pH readings matched spatially to waterbody and then to subwatershed'
       )
```

```{r visualize_number_calcium_data_points_per_subw}
watershed |> 
  dplyr::left_join(
    calc_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = log(n))) + 
  labs(title = 'Calcium (dissolved and total) Readings from EMS across BC',
       subtitle = paste0(nrow(calc_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(calc)/total_calc),1),'% of ~',round(total_calc/1000,0),'K points matched a waterbody'),
       fill = 'Natural log\nof number of \ndata points',
       caption = 'Calcium readings matched spatially to waterbody and then to subwatershed'
       )
```

## Mean Values 

```{r mean_pH_value_by_subw}
mean_pH_by_subw = pH |> 
  sf::st_drop_geometry() |> 
  dplyr::group_by(WATERSHED_) |> 
  dplyr::summarise(mean_pH = mean(RESULT, na.rm=T),
                   standard_deviation_pH = sd(RESULT, na.rm=T),
                   number_pH_data_points = n())

watershed |> 
  dplyr::left_join(
    mean_pH_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = mean_pH)) + 
  labs(title = 'pH Readings from EMS across BC',
       subtitle = paste0(nrow(mean_pH_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(pH)/total_pH),1),'% of ~',round(total_pH/1000,0),'K points matched some watershed'),
       fill = 'Mean pH Value',
       caption = 'pH readings matched spatially to subwatershed \ngroup directly, skipping match with waterbodies'
       )
```

```{r mean_calcium_value_by_subw}
mean_calc_by_subw = calc |> 
  sf::st_drop_geometry() |> 
  dplyr::group_by(WATERSHED_) |> 
  dplyr::summarise(mean_calcium = mean(RESULT, na.rm=T),
                   standard_deviation_calcium = sd(RESULT, na.rm=T),
                   number_calcium_data_points = n())

watershed |> 
  dplyr::left_join(
    mean_calc_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = mean_calcium)) + 
  labs(title = 'Calcium (dissolved and total) Readings from EMS across BC',
       subtitle = paste0(nrow(mean_calc_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(calc)/total_calc),1),'% of ~',round(total_calc/1000,0),'K points matched some watershed'),
       fill = 'Mean Calcium Value',
       caption = 'calcium readings matched spatially to subwatershed \ngroup directly, skipping match with waterbodies'
       ) + 
  scale_fill_fermenter(palette = 'RdYlGn')
```

```{r pH_in_ideal_ZQ_conditions}
pH_within_ideal_range = mean_pH_by_subw |>
  dplyr::mutate(
    ideal_conditions_ZQ = mean_pH >= 7 & mean_pH <= 9.6,
    ideal_conditions_Golden = mean_pH >= 6 & mean_pH <= 9
  )
```

```{r get_temperatures_per_subw}
#Grab the temperature data from the worldclim predicted temperature raster we read in at the top of the script.
watershed = watershed %>% 
  mutate(temp_mean = terra::extract(tempr, vect(watershed %>% st_transform(crs = 4326)), fun = mean, na.rm=T)[,2]) |> 
  mutate(temp_max = terra::extract(tempr, vect(watershed %>% st_transform(crs = 4326)), fun = max, na.rm=T)[,2])
  
watershed = watershed %>%  
  mutate(temp_group_zqm = case_when(
    temp_mean < temp_limit_zqm ~ paste0("< ", temp_limit_zqm),
    temp_mean >=  temp_limit_zqm ~ paste0(">= ", temp_limit_zqm),
    T ~ "NA"
  )) |> 
  mutate(temp_group_golden = case_when(
    temp_max < temp_limit_golden ~ paste0("< ", temp_limit_golden),
    temp_max >=  temp_limit_golden ~ paste0(">= ", temp_limit_golden),
    T ~ "NA"
  ))
```

## Temperature Correction Maps {.tabset}

### Zebra-Quagga

```{r temp_correction_map_ZQM}
ggplot() +
  geom_sf(data = watershed, col = "black", aes(fill = temp_group_zqm)) +
  scale_fill_manual(values = c("< 10" = "red",
                               "> 10" = "transparent")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() +
  labs(fill = "Average Air Temperature (C)")
```

### Golden 
```{r temp_correction_map_Golden}
ggplot() +
  geom_sf(data = watershed, col = "black", aes(fill = temp_group_golden)) +
  scale_fill_manual(values = c("< 15" = "red",
                               "> 15" = "transparent")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() +
  labs(fill = "Maximum Air Temperature (C)")
```

```{r join_calc_and_pH_to_watershed}
watershed = watershed |> 
  dplyr::left_join(
    mean_calc_by_subw
  ) |> 
  dplyr::left_join(
    mean_pH_by_subw
  )
```

```{r bin_calcium_values}

calc_bin_levels_zqm = data.frame(
  levels = c('No Risk (< 8 mg/L)',
             'Low (8 to <15 mg/L)',
             'Moderate (15 to <30 mg/L)',
             'High (30+ mg/L)',
             'No Data'),
  number = c(0,1,2,3, 99)
)

calc_bin_levels_golden = data.frame(
  levels = c('Low (0 to 1 mg/L)',
             'Moderate (1 to 3 mg/L)',
             'High (3+ mg/L)',
             'No Data'),
  number = c(1,2,3,99)
)
  
watershed = watershed |> 
  dplyr::mutate(calc_bin_zqm = case_when(
    mean_calcium < 8 ~ 'No Risk (< 8 mg/L)',
    mean_calcium >= 8 & mean_calcium < 15 ~ 'Low (8 to <15 mg/L)',
    mean_calcium >= 15 & mean_calcium < 30 ~ 'Moderate (15 to <30 mg/L)',
    mean_calcium >= 30 ~ 'High (30+ mg/L)',
    is.na(mean_calcium) ~ "No Data"
  )) |> 
  mutate(calc_bin_zqm = factor(calc_bin_zqm, levels = calc_bin_levels_zqm$levels)) |>
  mutate(calc_bin_n_zqm = case_when(
    calc_bin_zqm == 'No Risk (< 8 mg/L)' ~ 0,
    calc_bin_zqm == 'Low (8 to <15 mg/L)' ~ 1,
    calc_bin_zqm == 'Moderate (15 to <30 mg/L)' ~ 2,
    calc_bin_zqm == 'High (30+ mg/L)' ~ 3,
    calc_bin_zqm == 'No Data' ~ 99 
  )) |> 
  dplyr::mutate(calc_bin_golden = case_when(
    mean_calcium < 1 & mean_calcium < 15 ~ 'Low (0 to 1 mg/L)',
    mean_calcium >= 1 & mean_calcium < 3 ~ 'Moderate (1 to 3 mg/L)',
    mean_calcium >= 3 ~ 'High (3+ mg/L)',
    is.na(mean_calcium) ~ "No Data"
  )) |> 
  mutate(calc_bin_golden = factor(calc_bin_golden, levels = calc_bin_levels_golden$levels)) |> 
  mutate(calc_bin_n_golden = case_when(
    calc_bin_golden == 'Low (0 to 1 mg/L)' ~ 1,
    calc_bin_golden == 'Moderate (1 to 3 mg/L)' ~ 2,
    calc_bin_golden == 'High (3+ mg/L)' ~ 3,
    calc_bin_golden == 'No Data' ~ 99 
  ))
```

# Results {.tabset}

## Mean Calcium Bins

```{r mean_calcium_binned_by_subwatershed, fig.width = 10}
zqm_plot = ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_zqm), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk (< 8 mg/L)" = "green",
                               "Low (8 to <15 mg/L)" = "yellow",
                               "Moderate (15 to <30 mg/L)" = "orange",
                               "High (30+ mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Zebra Quagga Mussel Risk Estimates',
       subtitle = paste0(nrow(watershed[watershed$calc_bin_zqm != 'No Data',]), ' of 246 subwatersheds with data'))

golden_plot = ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_golden), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "Low (0 to 1 mg/L)" = "yellow",
                               "Moderate (1 to 3 mg/L)" = "orange",
                               "High (3+ mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Golden Mussel Risk Estimates',
       subtitle = paste0(nrow(watershed[watershed$calc_bin_zqm != 'No Data',]), ' of 246 subwatersheds with data'))

zqm_plot + golden_plot
```

```{r apply_pH_ranges_to_mean_calcium_risk_levels}
watershed = watershed |> 
  dplyr::left_join(
    pH_within_ideal_range
  ) |> 
  # Note: some watersheds did not receive qualifications of pH suitability for ZQM and Golden Mussel
  dplyr::mutate(
    calc_bin_n_zqm_pH_adj = ifelse(
      !ideal_conditions_ZQ & !is.na(ideal_conditions_ZQ) & calc_bin_n_zqm != 0 & calc_bin_n_zqm != 99, 
      calc_bin_n_zqm - 1, 
      calc_bin_n_zqm)
  ) |> 
  dplyr::mutate(
    calc_bin_n_golden_pH_adj = ifelse(
      !ideal_conditions_Golden & !is.na(ideal_conditions_Golden) & calc_bin_n_golden != 1 & calc_bin_n_golden != 99, 
      calc_bin_n_golden - 1, 
      calc_bin_n_golden)
  )
```

## Mean Calcium Bins, pH and Temperature Adjusted

Estimated Risk Levels based on available dissolved calcium data were reduced by 1 risk level in watersheds found to have non-suitable average pH and/or temperature. 
For Zebra / Quagga Mussels, unsuitable pH was < 7 or > 9.6, and temperatures were deemed unsuitable in any watersheds with annual mean temperatures under 10 degrees C. For Golden Mussel, unsuitable pH was < 6 or > 9, and temperatures were deemed unsuitable in any watersheds where the maximum annual temperature never goes above 15 degrees C.

```{r fig.width = 10}
# Apply temperature correction factor to risk estimates.
watershed = watershed |>
  mutate(calc_bin_n_zqm_pH_temp_adj = ifelse(
    temp_group_zqm == "< 10" & calc_bin_n_zqm_pH_adj != 0 & calc_bin_n_zqm_pH_adj != 99,
    calc_bin_n_zqm_pH_adj - 1,
    calc_bin_n_zqm_pH_adj
  )) |> 
  mutate(calc_bin_n_golden_pH_temp_adj = ifelse(
    temp_group_golden == "< 15" & calc_bin_n_golden_pH_adj != 1 & calc_bin_n_zqm_pH_temp_adj != 99,
    calc_bin_n_golden_pH_adj - 1,
    calc_bin_n_golden_pH_adj
  )) 

# Add labels for map.
watershed = watershed |> 
  dplyr::left_join(
    calc_bin_levels_zqm |> 
      dplyr::rename(calc_bin_n_zqm_pH_temp_adj = number,
                    calc_bin_zqm_pH_temp_adj = levels)
  ) |> 
  dplyr::left_join(
    calc_bin_levels_golden |> 
      dplyr::rename(calc_bin_n_golden_pH_temp_adj = number,
                    calc_bin_golden_pH_temp_adj = levels)
  ) |> 
  dplyr::mutate(calc_bin_zqm_pH_temp_adj = factor(
    calc_bin_zqm_pH_temp_adj,
    levels = c("No Data",
               "No Risk (< 8 mg/L)",
               "Low (8 to <15 mg/L)",
               "Moderate (15 to <30 mg/L)",
               "High (30+ mg/L)"))
    ) |> 
  dplyr::mutate(calc_bin_golden_pH_temp_adj = factor(
    calc_bin_golden_pH_temp_adj,
    levels = c("No Data",
               "Low (0 to 1 mg/L)",
               "Moderate (1 to 3 mg/L)",
               "High (3+ mg/L)"))
    )

zqm_plot = ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_zqm_pH_temp_adj), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "No Risk (< 8 mg/L)" = "green",
                               "Low (8 to <15 mg/L)" = "yellow",
                               "Moderate (15 to <30 mg/L)" = "orange",
                               "High (30+ mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() +
  labs(fill = 'Binned Calcium',
       title = 'Zebra / Quagga Mussel Risk Estimates',
       subtitle = 'pH and temperature adjusted') + 
  theme(legend.background = element_blank())

golden_plot = ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_golden_pH_temp_adj), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "Low (0 to 1 mg/L)" = "yellow",
                               "Moderate (1 to 3 mg/L)" = "orange",
                               "High (3+ mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() +
  labs(fill = 'Binned Calcium',
       title = 'Golden Mussel Risk Estimates',
       subtitle = 'pH and temperature adjusted')

my_patch = zqm_plot + golden_plot

my_patch

ggsave(filename = "output/ZQ_and_Golden_Mussels_calcium_suitability_comparisons.jpg", 
       my_patch,
       width = 10, height = 6, dpi = 300)

ggsave(filename = "J:/2 SCIENCE - Invasives/SPECIES/Golden mussel/ZQ_and_Golden_Mussels_calcium_suitability_comparisons.jpg", 
       my_patch,
       width = 10, height = 6, dpi = 300)
```

