---
title: "01_ZQMussel_Survival_Probability_Assessment"
author: "Chris Madsen"
date: "`r Sys.Date()`"
output: html_document
---

This script assesses the probability of survival of Zebra and Quagga Mussels in freshwater subwatersheds of BC, using environmental data (dissolved and total calcium, and water temperature from the environmental monitoring database). It is modelled after the 2017 DFO risk assessment of Canada for invasion of these species and Adam Sepulveda et al's more recent methodology. Quagga mussels are assessed first, a temperature correction factor is calculated, and this factor is then applied to the estimated risk factor of Zebra mussels.

# Sections {.tabset}

## Cleaning 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
# Load libraries
library(terra)
library(tidyverse)
library(lubridate)
library(sf)
library(ggspatial)
library(ggthemes)
library(bcdata)
```

```{r read_and_set_options}
my_opts = read_csv(paste0(here::here(),"/Options.csv")) %>% 
  as.data.frame()

my.year = my_opts$year

temp_limit = my_opts$zqm_risk_temp_limit
climate_model = my_opts$zqm_risk_climate_model

#Do we want to make monitoring location and water body level calcium shapefiles?
#interim_shapefile_output = F #F if no, T if yes.

setwd(paste0(my_opts$base_dir,"01_DataCleaning/"))
mywd = getwd()

#Create organizational folders in your working directory, if they don't exist.
#data...
if(!dir.exists(paste0(getwd(),"/data/"))) {
  dir.create(paste0(getwd(),"/data/"))
}

#shapefiles...
data.path = paste0(getwd(),"/data/")

# if(!dir.exists(paste0(data.path,"shapefiles"))) {
#   dir.create(paste0(data.path,"shapefiles"))
# }

#output...
if(!dir.exists(paste0(getwd(),"/output/"))) {
  dir.create(paste0(getwd(),"/output/"))
}
```

```{r load_in_shapefiles}
# BC shapefile
bc_shp = bcmaps::bc_bound() %>% 
  st_transform(crs = 4326)

# WorldClim future temperature raster data. 
if(my_opts$zqm_risk_climate_model == "ssp245"){
  tempr <- rast(paste0(data.path,"wc2.1_2.5m_bioc_MRI-ESM2-0_ssp245_2021-2040.tif"), lyrs = 10)
  tempr = mask(crop(tempr, bc_shp), vect(bc_shp))
} else if(my_opts$zqm_risk_climate_model == "ssp370"){
  tempr <- rast(paste0(data.path,"wc2.1_2.5m_bioc_MRI-ESM2-0_ssp370_2021-2040.tif"), lyrs = 10)
  tempr = mask(crop(tempr, bc_shp), vect(bc_shp))
}

#Environmental Monitoring System calcium data (total and dissolved)
#Load in the first dataset (last 2 years) - downloaded at the beginning of 2021, will need to be re-downloaded when run in the future!
df <- read_csv(paste0(data.path,"ems_sample_results_current_expanded.csv"), 
                col_types = cols(COLLECTION_END = col_datetime(format = "%Y%m%d%H%M%S"), 
                                 COLLECTION_START = col_datetime(format = "%Y%m%d%H%M%S"), 
                                 UPPER_DEPTH = col_integer(),
                                 LOWER_DEPTH = col_integer(),
                                 TIDE = col_integer(),
                                 AIR_FILTER_SIZE = col_integer(),
                                 AIR_FLOW_VOLUME = col_integer(),
                                 FLOW_UNIT = col_character(),
                                 COMPOSITE_ITEMS = col_integer(),
                                 CONTINUOUS_AVERAGE = col_integer(),
                                 CONTINUOUS_MAXIMUM = col_integer(),
                                 CONTINUOUS_MINIMUM = col_integer(),
                                 CONTINUOUS_UNIT_CODE = col_integer(),
                                 CONTINUOUS_DURATION = col_integer(),
                                 CONTINUOUS_DURATION_UNIT = col_character())) %>% 
  filter(is.na(LATITUDE)==F) %>% 
  mutate(UNIT = MDL_UNIT) |> 
  filter(PARAMETER_CODE %in% c("CA-T","CA-D","0004"))

#Get rid of columns 29 to 60, which we do not need.
df = df[,c(1:28)]

## Load in the mammoth dataset ###
mammoth_dataset = read_csv(paste0(data.path,"mammoth_dataset_filtered_for_total_and_dissolved_calcium_and_pH.csv"),

                           col_types = cols(COLLECTION_END = col_datetime(format = "%Y%m%d%H%M%S"),
                                            COLLECTION_START = col_datetime(format = "%Y%m%d%H%M%S"),
                                            UPPER_DEPTH = col_integer(),
                                            LOWER_DEPTH = col_integer(),
                                            TIDE = col_integer(),
                                            AIR_FILTER_SIZE = col_integer(),
                                            AIR_FLOW_VOLUME = col_integer(),
                                            FLOW_UNIT = col_character(),
                                            COMPOSITE_ITEMS = col_integer(),
                                            CONTINUOUS_AVERAGE = col_integer(),
                                            CONTINUOUS_MAXIMUM = col_integer(),
                                            CONTINUOUS_MINIMUM = col_integer(),
                                            CONTINUOUS_UNIT_CODE = col_integer(),
                                            CONTINUOUS_DURATION = col_integer(),
                                            CONTINUOUS_DURATION_UNIT = col_character()))

mammoth_dataset = mammoth_dataset[,c(1:28)]
```

```{r combine_data}
df = bind_rows(df, mammoth_dataset) %>% 
  dplyr::select(EMS_ID, MONITORING_LOCATION,LOCATION_TYPE,COLLECTION_START,COLLECTION_END,SAMPLING_AGENCY,COLLECTION_METHOD,PARAMETER,ANALYTICAL_METHOD,RESULT,UNIT,LONGITUDE,LATITUDE)
rm(mammoth_dataset)

hist(lubridate::year(df$COLLECTION_END)) #What's the distribution of data through time?
```

```{r filter_outliers_from_calcium_and_pH}
df_f = df |> 
  dplyr::filter(!is.na(RESULT)) |> 
  mutate(retain_point = case_when(
    PARAMETER == 'pH' & (RESULT <= 0 | RESULT > 16.5) ~ F,
    PARAMETER != 'pH' & (RESULT <= 0 | RESULT > 60) ~ F,
    T ~ T
  )) |> 
  dplyr::filter(retain_point)
```

```{r read_in_wb_and_subw_shapefiles}
#Read in the water body shapefiles
waterbodies = read_sf(paste0("data/summarized_bc_waterbodies_same_gnis_joined.gpkg"))
# Read in subwatershed (N = 246) shapefile
watershed = read_sf(paste0(my_opts$remote_spatial_data,"shared_data_sets/WatershedGroups_lowres.shp"))
```

```{r match_EMS_records_to_wb}
# ---------------
# We have a decision to make here: do we filter EMS data by whether or not it matches to a waterbody (either by name or spatially)?

# Calculate number of data points for pH and calcium
total_pH = nrow(df_f |> 
                  dplyr::filter(PARAMETER == 'pH'))

total_calc = nrow(df_f |> 
                  dplyr::filter(PARAMETER != 'pH'))

# Simple spatial match to waterbodies

df_f_match_to_wbs = df_f |>
  dplyr::filter(!is.na(LATITUDE)) |> 
  st_as_sf(coords = c('LONGITUDE','LATITUDE'),
           crs = 4326) |> 
  sf::st_transform(crs = 3005) |> 
  st_join(waterbodies)

df_f_match_to_wbs = df_f_match_to_wbs |>
  dplyr::filter(!is.na(WATERSH))

# If we match to waterbody, we keep about 198K data points.

# length(unique(df_f_match_to_wbs$WATERSH))

# # ---------------------
# # Simple spatial match to subwatershed
# 
# df_f_match_to_subw = df_f |>
#   dplyr::filter(!is.na(LATITUDE)) |> 
#   st_as_sf(coords = c('LONGITUDE','LATITUDE'),
#            crs = 4326) |> 
#   sf::st_transform(crs = 3005) |> 
#   st_join(watershed)
# 
# df_f_match_to_subw |> 
#   dplyr::filter(!is.na(WATERSHED_))
# 
# # If we match to subwatershed, we keep about 741K data points.
# 
# length(unique(df_f_match_to_subw$WATERSHED_))




# If we just want to match to watershed, can do it here.
# watershed = read_sf(paste0(my_opts$remote_spatial_data,"shared_data_sets/WatershedGroups_lowres.shp"))
# 
# watershed = watershed |> dplyr::select(watershed_name = WATERSHE_1,
#                                      watershed_number = WATERSHED_)

pH = df_f_match_to_wbs |> filter(PARAMETER == 'pH') |> dplyr::rename(WATERSHED_ = WATERSH)

calc = df_f_match_to_wbs |> filter(PARAMETER != 'pH') |> dplyr::rename(WATERSHED_ = WATERSH)

# pH_match = pH |> 
#   st_join(watershed) |> 
#   dplyr::filter(!is.na(watershed_name))
# 
# calc_match = calc |> 
#   st_join(watershed) |> 
#   dplyr::filter(!is.na(watershed_name))

# # Summarized number of data points for pH and calcium by subwatershed.
pH_by_subw = pH |>
  sf::st_drop_geometry() |>
  dplyr::count(WATERSHED_)

calc_by_subw = calc |>
  sf::st_drop_geometry() |>
  dplyr::count(WATERSHED_)

```

## Number of Data Points

```{r visualize_number_pH_data_points_per_subw}
# Add summaries to subwatershed; plot.
watershed |> 
  dplyr::left_join(
    pH_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = log(n))) + 
  labs(title = 'pH Readings from EMS across BC',
       subtitle = paste0(nrow(pH_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(pH)/total_pH),1),'% of ~',round(total_pH/1000,0),'K points matched a waterbody'),
       fill = 'Natural log\nof number of \ndata points',
       caption = 'pH readings matched spatially to waterbody and then to subwatershed'
       )

ggsave(filename = 'output/pH_EMS_records_spatially_matched_to_waterbodies.png',
       width = 8,
       height = 8)

```

```{r visualize_number_calcium_data_points_per_subw}
watershed |> 
  dplyr::left_join(
    calc_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = log(n))) + 
  labs(title = 'Calcium (dissolved and total) Readings from EMS across BC',
       subtitle = paste0(nrow(calc_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(calc)/total_calc),1),'% of ~',round(total_calc/1000,0),'K points matched a waterbody'),
       fill = 'Natural log\nof number of \ndata points',
       caption = 'Calcium readings matched spatially to waterbody and then to subwatershed'
       )

ggsave(filename = 'output/calc_EMS_records_spatially_matched_to_waterbodies.png',
       width = 8, 
       height = 8)
```

## Mean Values 

```{r mean_pH_value_by_subw}
mean_pH_by_subw = pH |> 
  sf::st_drop_geometry() |> 
  dplyr::group_by(WATERSHED_) |> 
  dplyr::summarise(mean_pH = mean(RESULT, na.rm=T),
                   standard_deviation_pH = sd(RESULT, na.rm=T),
                   number_pH_data_points = n())

watershed |> 
  dplyr::left_join(
    mean_pH_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = mean_pH)) + 
  labs(title = 'pH Readings from EMS across BC',
       subtitle = paste0(nrow(mean_pH_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(pH)/total_pH),1),'% of ~',round(total_pH/1000,0),'K points matched some watershed'),
       fill = 'Mean pH Value',
       caption = 'pH readings matched spatially to subwatershed \ngroup directly, skipping match with waterbodies'
       )

ggsave(filename = 'output/mean_pH_by_subwatershed_group.png',
       width = 8, height = 8)
```

```{r mean_calcium_value_by_subw}
mean_calc_by_subw = calc |> 
  sf::st_drop_geometry() |> 
  dplyr::group_by(WATERSHED_) |> 
  dplyr::summarise(mean_calcium = mean(RESULT, na.rm=T),
                   standard_deviation_calcium = sd(RESULT, na.rm=T),
                   number_calcium_data_points = n())

watershed |> 
  dplyr::left_join(
    mean_calc_by_subw
  ) |> 
  ggplot() +
  geom_sf(aes(fill = mean_calcium)) + 
  labs(title = 'Calcium (dissolved and total) Readings from EMS across BC',
       subtitle = paste0(nrow(mean_calc_by_subw),' of 246 subwatersheds have data; ',round(100*(nrow(calc)/total_calc),1),'% of ~',round(total_calc/1000,0),'K points matched some watershed'),
       fill = 'Mean Calcium Value',
       caption = 'calcium readings matched spatially to subwatershed \ngroup directly, skipping match with waterbodies'
       ) + 
  scale_fill_fermenter(palette = 'RdYlGn')

ggsave(filename = 'output/mean_calcium_by_subwatershed_group.png',
       width = 8, height = 8)
```

## pH Suitability 

```{r pH_in_ideal_ZQ_conditions}
pH_within_ideal_ZQ_range = mean_pH_by_subw |> 
  dplyr::mutate(ideal_conditions = mean_pH >= 7 & mean_pH <= 9.6)

watershed |> 
  dplyr::left_join(
    pH_within_ideal_ZQ_range
  ) |> 
  ggplot() +
  geom_sf(aes(fill = ideal_conditions)) + 
  labs(fill = 'Mean pH within \nideal ZQ Mussel \nrange(7 to 9.6)')

ggsave(filename = 'output/mean_pH_within_ideal_ZQ_Mussels_range_by_subwatershed.png',
       width = 8, height = 8)

openxlsx::write.xlsx(pH_within_ideal_ZQ_range, 'output/mean_pH_waterbody_matched_within_ideal_ZQ_Mussels_range_by_subwatershed.xlsx')

```

```{r get_temperatures_per_subw}
#Grab the temperature data from the worldclim predicted temperature raster we read in at the top of the script.
watershed = watershed %>% 
  mutate(temp = terra::extract(tempr, vect(watershed %>% st_transform(crs = 4326)), fun = mean, na.rm=T)[,2])

watershed = watershed %>%  
  mutate(temp_group = case_when(
    temp < temp_limit ~ paste0("< ", temp_limit),
    temp >=  temp_limit ~ paste0(">= ", temp_limit),
    T ~ "NA"
  ))
```

## Temperature Correction Map
```{r temp_correction_map}
my_plot = ggplot() +
  geom_sf(data = watershed, col = "black", aes(fill = temp_group)) +
  scale_fill_manual(values = c("< 10" = "red",
                               "> 10" = "transparent")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() +
  labs(fill = "Air Temperature (C)")

my_plot

ggsave(filename = paste0(mywd,"/output/TemperatureCorrection_",temp_limit,"_",climate_model,"_RiskMap.png"), plot = my_plot, dpi = 150, width = 10, height = 7)
```

```{r join_calc_and_pH_to_watershed}
watershed = watershed |> 
  dplyr::left_join(
    mean_calc_by_subw
  ) |> 
  dplyr::left_join(
    mean_pH_by_subw
  )
```

```{r bin_calcium_values}

calc_bin_levels = data.frame(
  levels = c('Minimal (< 8 mg/L)',
             'Moderate (8 to 20 mg/L)',
             'High (> 20 mg/L)',
             'No Data'),
  number = c(1,2,3, 99)
)
  
watershed = watershed |> 
  dplyr::mutate(calc_bin = case_when(
    mean_calcium < 8 ~ 'Minimal (< 8 mg/L)',
    mean_calcium >= 8 & mean_calcium < 20 ~ 'Moderate (8 to 20 mg/L)',
    mean_calcium >= 20 ~ 'High (> 20 mg/L)',
    is.na(mean_calcium) ~ "No Data"
  )) |> 
  mutate(calc_bin = factor(calc_bin, levels = calc_bin_levels$levels)) |> 
  mutate(calc_bin_n = case_when(
    calc_bin == 'Minimal (< 8 mg/L)' ~ 1,
    calc_bin == 'Moderate (8 to 20 mg/L)' ~ 2,
    calc_bin == 'High (> 20 mg/L)' ~ 3,
    calc_bin == 'No Data' ~ 99 
  ))
```

# Results {.tabset}

## Mean Calcium Bins

```{r mean_calcium_binned_by_subwatershed}
ggplot(watershed) +
  geom_sf(aes(fill = calc_bin), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "Minimal (< 8 mg/L)" = "green",
                               "Moderate (8 to 20 mg/L)" = "yellow",
                               "High (> 20 mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Binned Calcium at the Watershed Group scale',
       subtitle = paste0(nrow(watershed[watershed$calc_bin != 'No Data',]), ' of 246 subwatersheds with data'))

ggsave(filename = 'output/Binned_Mean_Calcium_by_subwatershed.png',
       width = 8, height = 8)
```

## Mean Calcium Bins, pH Adjusted

```{r mean_calcium_binned_and_reduced_by_pH_by_subwatershed}

# Identify watersheds with non-ideal pH
non_ideal_pH_subw = pH_within_ideal_ZQ_range |> 
  dplyr::filter(!ideal_conditions) |> 
  pull(WATERSHED_)

watershed = watershed |> 
  mutate(calc_bin_n_pH_adj = case_when(
    WATERSHED_ %in% non_ideal_pH_subw & calc_bin_n != 1 & calc_bin_n != 99 ~ calc_bin_n - 1,
    T ~ calc_bin_n
  ))

watershed = watershed |> 
  left_join(calc_bin_levels |> 
              rename(calc_bin_n_pH_adj = number)) |> 
  mutate(calc_bin_pH_adj = levels) |> 
  dplyr::select(-levels) |> 
  mutate(calc_bin_pH_adj = factor(calc_bin_pH_adj, levels = calc_bin_levels$levels))

ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_pH_adj), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "Minimal (< 8 mg/L)" = "green",
                               "Moderate (8 to 20 mg/L)" = "yellow",
                               "High (> 20 mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Binned Calcium (Adjusted by pH suitability) at the Watershed Group scale',
       subtitle = paste0(nrow(watershed[watershed$calc_bin != 'No Data',]), ' of 246 with data; unsuitable pH (<7 or >9.6) reduced risk by one level (if not already "Minimal")'))

ggsave(filename = 'output/Binned_Mean_Calcium_pH_Adjusted_by_subwatershed.png',
       width = 8, height = 8)
```

## Mean Calcium Bins, pH and Temperature Adjusted

```{r mean_calcium_binned_and_reduced_by_pH_reduced_by_low_temps_by_subwatershed}

# Identify watersheds with non-ideal pH
low_temp_watersheds = watershed |> 
  dplyr::filter(temp_group == '< 10') |> 
  pull(WATERSHED_)

watershed = watershed |> 
    mutate(calc_bin_n_pH_temp_adj = case_when(
        WATERSHED_ %in% low_temp_watersheds & calc_bin_n_pH_adj != 1 & calc_bin_n_pH_adj != 99 ~ calc_bin_n_pH_adj - 1,
        T ~ calc_bin_n_pH_adj
    ))

watershed = watershed |> 
  left_join(calc_bin_levels |> 
              rename(calc_bin_n_pH_temp_adj = number)) |> 
  mutate(calc_bin_pH_temp_adj = levels) |> 
  dplyr::select(-levels) |> 
  mutate(calc_bin_pH_temp_adj = factor(calc_bin_pH_temp_adj, levels = calc_bin_levels$levels))

ggplot(watershed) +
  geom_sf(aes(fill = calc_bin_pH_temp_adj), col = "black") +
  scale_fill_manual(values = c("No Data" = "grey",
                               "Minimal (< 8 mg/L)" = "green",
                               "Moderate (8 to 20 mg/L)" = "yellow",
                               "High (> 20 mg/L)" = "red")) +
  annotation_scale(location = "br", width_hint = 0.5) +
  theme_map() + 
  labs(fill = 'Binned Calcium',
       title = 'Binned Calcium (Adjusted by pH and temperature suitability) at the Watershed Group scale',
       subtitle = paste0(nrow(watershed[watershed$calc_bin != 'No Data',]), ' of 246 with data; unsuitable pH (<7 or >9.6) reduced risk by one level (if not already "Minimal")'))

ggsave(filename = 'output/Binned_Mean_Calcium_pH_and_Temp_Adjusted_by_subwatershed.png',
       width = 8, height = 8)
```

```{r write_sf_to_disk}
write_sf(watershed,
         paste0("output/ZQM_RiskAssessment_",my_opts$zqm_risk_climate_model,"_",
                temp_limit,"C_threshold_pH_Adjustment_Subwatershed.gpkg"))

# ### Produce excel files and shapefiles
# watdat = watdat %>% 
#   mutate(Quagga_Risk = "NA") %>%
#   mutate(Quagga_Risk = replace(Quagga_Risk, bin == 1, "Minimal (less than 8 mg/L)"),
#          Quagga_Risk = replace(Quagga_Risk, bin == 2, "Moderate (between 8 and 20 mg/L)"),
#          Quagga_Risk = replace(Quagga_Risk, bin == 3, "High (greater than 20 mg/L)"),
#          Zebra_Risk = "NA",
#          Zebra_Risk = replace(Zebra_Risk, bin_bumped == 1, "Minimal (less than 8 mg/L)"),
#          Zebra_Risk = replace(Zebra_Risk, bin_bumped == 2, "Moderate (between 8 and 20 mg/L)"),
#          Zebra_Risk = replace(Zebra_Risk, bin_bumped == 3, "High (greater than 20 mg/L)"))
# 
# write_sf(watdat, 
#          paste0(mywd,"/output/ZQM_RiskAssessment_",my_opts$zqm_risk_climate_model,"_",
#                 temp_limit,"C_threshold_Subwatershed.gpkg"))
```
