

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse)
library(openxlsx)
library(dplyr)
library(data.table)
library(kableExtra)


my_opts = read_csv(paste0(here::here(),"/Options.csv"))


my.year = 2026
zq<- read.xlsx(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/waterbodies with risk estimates post filtering_",my.year,".xlsx"))

golden<-read.xlsx(paste0(my_opts$zqm_operations_data_folder,"Lake Monitoring/",my.year,"/Prioritization model/waterbodies with risk estimates post filtering_",my.year,"_golden_mussel.xlsx"))

# waterb = read_sf(paste0(my_opts$base_dir,"03_PrioritizationModel/data/waterb_with_data.gpkg"))

```

# Differences in Model Output {.tabset}

First, waterbodies that are on the final watebody list for Zebra Quagga Mussel. Then, those that are on the list for Golden Mussel, but NOT for Zebra Quagga mussel

## Zebra Quaqqa not on Golden Mussel List

pH range for Zebra Mussel: 7 to 9.6 

Calcium for Zebra Mussel: 8 to 30+ mg/l

Temperature: >10 degrees celcius


```{r check_zq}

zq_not_in_golden_full <- zq %>%
  anti_join(
    golden,
    by = c("WatershedName", "GNIS_NA")
  )

DT::datatable(zq_not_in_golden_full,
              options = list(autoWidth = FALSE,
    scrollX = TRUE,
    pageLength = 10,
    lengthChange = FALSE))


write.xlsx(zq_not_in_golden_full, "./output/zq_not_on_golden.xlsx")

```

## Golden Mussel not on Zebra Quagga List

pH range for Zebra Mussel: 6 to 9

Calcium for Zebra Mussel: 1 to 12+ mg/l

Temperature: >15 degrees celcius

```{r check_golden}

golden_not_zq_full <- golden %>%
  anti_join(
    zq,
    by = c("WatershedName", "GNIS_NA")
  )

DT::datatable(golden_not_zq_full,
              options = list(autoWidth = FALSE,
    scrollX = TRUE,
    pageLength = 10,
    lengthChange = FALSE))


write.xlsx(golden_not_zq_full, "./output/golden_not_on_zq.xlsx")

```























